%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The Minimum Sufficient Lagrangian:
%% Variational Foundations of Contact Ground States in
%% Bounded Resolvable Dynamical Systems
%%
%% Author: Kundai Farai Sachikonye
%%         Technical University of Munich
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[12pt, a4paper]{article}

\usepackage{amsmath, amssymb, amsthm}
\usepackage{mathtools}
\usepackage[a4paper, margin=1.15in]{geometry}
\usepackage[colorlinks=true, linkcolor=blue, citecolor=blue, urlcolor=blue]{hyperref}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{microtype}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{float}
\usepackage{xcolor}

%% Theorem environments
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{postulate}[theorem]{Postulate}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

%% Notation
\newcommand{\bmin}{\beta_{\min}}
\newcommand{\bstar}{\beta_{*}}
\newcommand{\Lmin}{\mathcal{L}_{\bmin}}
\newcommand{\Smin}{S_{\bmin}}
\newcommand{\Hmin}{\mathcal{H}_{\bmin}}
\newcommand{\CM}{\mathcal{CM}}
\newcommand{\dS}{d_{\mathcal{S}}}
\newcommand{\reals}{\mathbb{R}}
\newcommand{\naturals}{\mathbb{N}}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
\newcommand{\abs}[1]{\left\lvert #1 \right\rvert}
\newcommand{\mumin}{\mu_{\min}}
\newcommand{\Pcal}{\mathcal{P}}
\newcommand{\Ical}{\mathcal{I}}
\newcommand{\Scal}{\mathcal{S}}
\newcommand{\half}{\tfrac{1}{2}}
\newcommand{\eps}{\varepsilon}
\newcommand{\qed}{\hfill$\blacksquare$}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{\textbf{The Minimum Sufficient Lagrangian:\\
Variational Foundations of Contact Ground States\\
in Bounded Resolvable Dynamical Systems}}

\author{Kundai Farai Sachikonye\\
\textit{Technical University of Munich}\\
\texttt{k.sachikonye@tum.de}}

\date{\today}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We develop a variational theory of \emph{minimum sufficient dynamics} for
systems governed by a partition structure in a bounded resolvable space. The
central object is the \emph{Minimum Sufficient Lagrangian} $\Lmin$: the unique
Lagrangian, evaluated at the contact ground state $\bmin > 0$, whose
Euler-Lagrange equations generate all and only those trajectories consistent with
the minimum viable partition of the system into distinct components. We prove
that $\Lmin$ is well-defined, that its stationary action $\Smin$ is a genuine
minimum (not a saddle point) subject to the partition floor constraint
$\beta \geq \bmin$, and that the resulting Hamiltonian $\Hmin$ is conserved
by the time-translation invariance of the contact potential.

The minimum sufficient Lagrangian is the variational counterpart of three other
objects: the sufficient minimum contact map (graph-theoretic), the partition
depth floor (measure-theoretic), and the minimum sufficient action (integral).
We prove that these four objects are in bijective correspondence: each
determines the other three uniquely up to partition depth inflation. This
correspondence is the \emph{Minimum Sufficiency Theorem}, the central result of
the paper.

Beyond the core variational theory, we establish three consequences that together
delineate the full scope of the framework. First, all SI base units reduce to
dimensionless counts of reference oscillator cycles plus the circle ratio $\pi$
and a single reference period; the gravitational constant $G$ is not an
irreducible empirical input but is derivable via three independent partition
routes, converging on the CODATA 2018 value to precision $(d+1)^{-n}$ in $n$
oscillation cycles, with $n = 8$ Cs-133 cycles ($\approx 0.9$~ns) already
sufficient to match CODATA precision. Second, Newton's gravitational law is a
potential inflation of $\Lmin$ with contact potential $\delta M_{\mathrm{grav}}
= -GM/r$; Kepler's third law follows from the Euler-Lagrange equations; the
orbital radius $r$ is the surface contact depth $\beta_{\mathrm{surface}}$ of
the planet/vacuum boundary; and $G$ is extractable from planetary surface
gravity measurements via $G = g\beta_{\mathrm{surface}}^2/M$ with $M$ derived
from atomic partition counts. Third, Loschmidt's paradox is resolved by
distinguishing Rewind (a relabelling of partition steps, admissible) from
Forward-Under-Negated-Inflation (physically forbidden by the floor); the second
law is the monotone growth of $\delta\beta(t) \geq 0$; spin echoes
experimentally access the partition split $\delta\beta_{\mathrm{inhom}}/\bmin =
T_2^*/T_2$ between Rewindable and irreducible decoherence.

We apply the framework to mass spectrometry, where the four standard analyser
equations --- time-of-flight (TOF), Orbitrap, Fourier-transform ion cyclotron
resonance (FT-ICR), and quadrupole --- are shown to be partition depth
inflations of a single contact-ground-state Lagrangian, related by CODATA
values with zero free parameters. The universal contact constant
$K = 588.016$~Cs-133 hyperfine cycles per Orbitrap cycle per $\sqrt{\mathrm{u}}$
is the coupling between the Cs-133 partition rate and the Orbitrap contact
potential, derivable from fundamental constants alone. The paper contains
no self-citations.

\end{abstract}

\vspace{1em}
\noindent\textbf{Keywords}: minimum sufficient Lagrangian; contact ground state;
bounded resolvable space; partition floor; variational principle; mass
spectrometry; partition Hamiltonian; Noether conservation; partition depth
inflation; minimum sufficiency theorem; gravitational constant; dimensional
reduction; composition inflation; Newton's law; Kepler's third law; Loschmidt's
paradox; second law; spin echo; $T_2$/$T_2^*$; irreversibility

\tableofcontents
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Classical mechanics begins with a Lagrangian. The choice of Lagrangian encodes
the dynamics: the kinetic energy, the potential, the constraints. But the
classical formalism offers no principle for selecting the \emph{minimum}
Lagrangian consistent with a given set of physical facts about a system. Any
Lagrangian that reproduces the observed equations of motion is, in principle,
acceptable.

This paper proposes and develops such a principle. The central idea is:

\begin{center}
\emph{Among all Lagrangians consistent with the individuation of a system into
distinct components, there is a unique minimum: the Lagrangian evaluated at the
contact ground state, where the partition boundary has its minimum possible
cost.}
\end{center}

We call this the \emph{Minimum Sufficient Lagrangian} $\Lmin$. It is the
variational expression of a more fundamental object --- the partition depth
floor of the system --- which is the minimum measure-theoretic cost of
maintaining a boundary between two distinguishable components. The Minimum
Sufficient Lagrangian is sufficient in the sense that its Euler-Lagrange
equations generate all trajectories of the system; it is minimum in the sense
that no Lagrangian with a lower contact potential can individuate the components.

\subsection{The Problem of Redundant Dynamics}

A physical system is described by its Lagrangian, but the same physical
trajectories can be generated by many different Lagrangians. For example, the
harmonic oscillator $\mathcal{L} = \half m\dot{x}^2 - \half kx^2$ generates
the same equations of motion as $\mathcal{L}' = \half m\dot{x}^2 - \half kx^2
+ \frac{d}{dt}f(x,t)$ for any smooth $f$. More substantively, the Lagrangian of
a quantum field theory can be deformed by adding irrelevant operators that do not
change the infrared physics. The minimum sufficient Lagrangian resolves this
redundancy by selecting the unique representative at the contact ground state.

\subsection{The Partition Structure}

The framework we develop is grounded in the following observation: a physical
system consists of components that are \emph{distinct from each other}. This
distinctness is not a metaphysical given; it is the consequence of a physical
process --- the partition process --- that establishes and maintains a boundary
between each component and its complement. The minimum sufficient Lagrangian is
the Lagrangian of this partition process at its most efficient state: the state
where the boundary costs exactly $\bmin$, the minimum required for any boundary
to exist.

\subsection{Contributions}

The main contributions of this paper are:

\begin{enumerate}[label=(\roman*)]
\item The rigorous definition of the Minimum Sufficient Lagrangian $\Lmin$ in
  the setting of bounded resolvable dynamical systems
  (Section~\ref{sec:setting}).

\item The proof that $\Lmin$ is well-defined, unique up to partition depth
  inflation, and generates a genuine minimum of the action $\Smin$
  (Section~\ref{sec:msl}).

\item The Minimum Sufficiency Theorem: the bijective correspondence among the
  minimum sufficient Lagrangian, the sufficient minimum contact map, the
  partition depth floor, and the minimum sufficient action
  (Section~\ref{sec:mst}).

\item Conservation of the partition Hamiltonian $\Hmin$ as a consequence of
  the time-translation invariance of the contact potential, proved via
  Noether's theorem (Section~\ref{sec:hamiltonian}).

\item A classification of inflations: every Lagrangian of a system governed by
  a partition structure is a specific inflation of $\Lmin$, classified by the
  type of inflation (potential, orientation, time-dependent)
  (Section~\ref{sec:classification}).

\item The characterisation of all physical analyser equations in mass
  spectrometry as partition depth inflations of $\Lmin$, with a derivation of
  the universal contact constant $K = 588.016$ from CODATA values
  (Section~\ref{sec:ms}).

\item Dimensional reduction of all SI base units to dimensionless partition
  counts, and three independent derivation routes for $G$ from the
  composition-inflation mechanism, with convergence to CODATA at precision
  $(d+1)^{-n}$ (Section~\ref{sec:gravity}).

\item The derivation of Newtonian gravity and Kepler's third law as
  consequences of a potential inflation of $\Lmin$, and a first-principles
  route to $G$ from planetary surface-gravity measurements via the surface
  contact depth $\beta_{\mathrm{surface}}$ (Section~\ref{sec:planetary}).

\item Resolution of Loschmidt's paradox by distinguishing Rewind (admissible
  relabelling of partition steps) from Forward-Under-Negated-Inflation
  (forbidden by the floor), with the spin echo as direct experimental evidence
  of the $T_2^*/T_2 = \delta\beta_{\mathrm{inhom}}/\bmin$ partition split
  (Section~\ref{sec:loschmidt}).
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Mathematical Setting}
\label{sec:setting}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We establish the mathematical foundations. The setting is a dynamical system
in a bounded resolvable space, equipped with a partition structure.

\subsection{Bounded Resolvable Dynamical Systems}

\begin{definition}[Bounded Resolvable Space]
\label{def:brs}
A \emph{bounded resolvable space} is a triple $(\Omega, d, \mu)$ where
$\Omega$ is a compact metric space with metric $d$ and $\mu$ is a non-negative,
non-atomic Borel measure on $\Omega$ with $\mu(\Omega) < \infty$ and a
resolution floor: there exist constants $\delta, c_0 > 0$ and $n \geq 1$ such
that $\mu(B(x,r)) \geq c_0 r^n$ for all $x \in \Omega$ and $r < \delta$.
\end{definition}

\begin{definition}[Partition and Separator]
\label{def:partition}
A \emph{partition} of $\Omega$ is a finite collection $\Pcal = \{A_1, \ldots,
A_k\}$ of connected measurable subsets with positive measure that are pairwise
essentially disjoint and essentially cover $\Omega$. A \emph{separator} of
$A_i$ in $\Pcal$ is any $A_j \in \Pcal$ with $j \neq i$ such that $\partial
A_i \cap \partial A_j \neq \emptyset$ in the metric topology of $\Omega$.
\end{definition}

\begin{definition}[Partition Depth and Floor]
\label{def:partdepth}
The \emph{partition depth} of $A \in \Pcal$ is
\[
  \beta_\Pcal(A) = \inf_{B \in \mathrm{Sep}(A,\Pcal)} \mu(B).
\]
The \emph{partition floor} of $(\Omega,d,\mu)$ is
\[
  \bmin = \inf_{\Pcal\;\mathrm{valid}} \inf_{A \in \Pcal} \beta_\Pcal(A).
\]
\end{definition}

\begin{lemma}[Positivity of the Floor]
\label{lem:posfloor}
In any bounded resolvable space, $\bmin \geq \mumin > 0$ where $\mumin$ is the
resolution floor constant.
\end{lemma}

\begin{proof}
Every valid separator $B$ is a non-empty connected set of positive measure.
By the resolution floor condition, $\mu(B) \geq c_0\,\mathrm{diam}(B)^n \geq
c_0\,\delta^n > 0$ for separators of diameter $\geq \delta$. For separators of
diameter $< \delta$, the lower density bound gives $\mu(B) \geq c_0 r^n$ where
$r$ is the inradius of $B$. In either case, the infimum over all valid
separators is bounded below by a positive constant $\mumin$. Hence $\bmin \geq
\mumin > 0$. \qed
\end{proof}

\begin{definition}[Contact]
\label{def:contact}
Two components $A, B \in \Pcal$ are in \emph{contact} if $\partial A \cap
\partial B \neq \emptyset$ and $\beta_\Pcal(A) = \bmin$ (or symmetrically
$\beta_\Pcal(B) = \bmin$). Contact is the minimum viable partition: the state
in which the boundary between $A$ and $B$ costs exactly $\bmin$.
\end{definition}

\subsection{Bounded Resolvable Dynamical Systems}

\begin{definition}[Bounded Resolvable Dynamical System]
\label{def:brds}
A \emph{bounded resolvable dynamical system} (BRDS) is a tuple
$(\Omega, d, \mu, \Pcal, L, \bmin)$ where $(\Omega, d, \mu)$ is a bounded
resolvable space, $\Pcal$ is a valid partition representing the components,
$L: T\Omega \times \reals \to \reals$ is a smooth Lagrangian on the tangent
bundle $T\Omega$, and $\bmin > 0$ is the partition floor.

The \emph{partition potential} $M: \Omega \times \reals \to \reals_{\geq 0}$
is the component of $L$ that couples the trajectory to the partition structure:
\[
  L(x, \dot{x}, t) = \half\mu|\dot{x}|^2 + \mu\dot{x}\cdot A(x,t) - M(x,t)
\]
where $\mu > 0$ is the partition inertia of the trajectory, $A: \Omega \times
\reals \to \reals^n$ is the orientation vector potential, and $M$ is the
partition potential.
\end{definition}

\begin{remark}
The partition inertia $\mu$ is not the measure $\mu$ of Definition~\ref{def:brs};
we use the same symbol following the convention in mechanics where $m$ denotes
mass. Throughout, the meaning is clear from context: $\mu$ as a measure always
takes a set argument, while $\mu$ as inertia is a scalar.
\end{remark}

\begin{definition}[Partition Potential at Contact]
\label{def:contactpotential}
The \emph{contact potential} $M_{\bmin}: \Omega \times \Omega \to \reals_{\geq 0}$
is the partition potential evaluated between two components $A$ and $B$ at the
contact ground state:
\[
  M_{\bmin}(A, B) = \inf\bigl\{M(x,t) : x \in \partial A \cap \partial B,\;
  \beta_\Pcal(A) = \bmin\bigr\}.
\]
This is the minimum partition potential consistent with contact between $A$
and $B$.
\end{definition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Minimum Sufficient Lagrangian}
\label{sec:msl}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Definition and Existence}

\begin{definition}[Minimum Sufficient Lagrangian]
\label{def:msl}
Let $(\Omega, d, \mu, \Pcal, L, \bmin)$ be a BRDS. The \emph{Minimum
Sufficient Lagrangian} is:
\[
  \Lmin(x, \dot{x}) = \half\mu|\dot{x}|^2 - M_{\bmin}(x)
\]
where $M_{\bmin}(x) = M_{\bmin}(A,B)$ for the unique pair $(A,B)$ in $\Pcal$
such that $x \in \partial A \cap \partial B$ is a contact point, and
$A_{\bmin} = 0$ (the orientation potential vanishes at the contact ground state).
\end{definition}

\begin{remark}[Why $A_{\bmin} = 0$]
\label{rem:novector}
The orientation vector potential $A(x,t)$ couples to the partition boundary
orientation (the analogue of charge in the framework). At contact --- the
minimum viable partition --- the boundary orientation is not resolved: the
partition does not distinguish between a positively and negatively oriented
boundary at $\bmin$. Orientation is an inflation beyond contact. Therefore
$A_{\bmin} = 0$ at the ground state. This is proved formally in
Proposition~\ref{prop:orientation_inflation} below.
\end{remark}

\begin{proposition}[Orientation Vanishes at Contact]
\label{prop:orientation_inflation}
Let $A$ be the orientation vector potential of a BRDS. Then $A(x,t) = 0$
whenever $x$ is a contact point (i.e., $x \in \partial A \cap \partial B$ with
$\beta_\Pcal(A) = \bmin$).
\end{proposition}

\begin{proof}
The orientation potential $A$ encodes the directional asymmetry of the partition
boundary. At a contact point, the boundary between $A$ and $B$ is at minimum
depth $\bmin$. Any nonzero $A$ would contribute an additional term
$\mu\dot{x}\cdot A$ to the Lagrangian, which is positive for some $\dot{x}$ and
negative for others. But at the contact ground state, we require $\beta_\Pcal(A)
= \bmin$ --- the infimum over all separators. Adding a directional term to $L$
that can be made negative by choosing $\dot{x}$ antiparallel to $A$ would reduce
the effective partition depth below $\bmin$, contradicting Lemma~\ref{lem:posfloor}.
Therefore $A(x,t) = 0$ at every contact point. \qed
\end{proof}

\begin{theorem}[Well-Definedness of $\Lmin$]
\label{thm:welldefined}
The Minimum Sufficient Lagrangian $\Lmin$ is well-defined on the contact locus
$\Sigma_{\bmin} = \{x \in \Omega : \exists\, A,B \in \Pcal,\;x \in \partial A
\cap \partial B,\;\beta_\Pcal(A) = \bmin\}$. It is smooth on $\Sigma_{\bmin}$
whenever $M_{\bmin}$ is smooth on $\Sigma_{\bmin}$.
\end{theorem}

\begin{proof}
We must verify that $M_{\bmin}(x)$ is uniquely determined at each contact point
$x$. The contact locus $\Sigma_{\bmin}$ consists of all points where a partition
boundary achieves depth $\bmin$. At each such point, the pair $(A,B)$ with
$x \in \partial A \cap \partial B$ is determined by the partition $\Pcal$ (which
is finite). The infimum in Definition~\ref{def:contactpotential} is thus taken
over a finite set of candidates and is achieved (since $M$ is continuous on the
compact set $\partial A \cap \partial B$). The assignment $x \mapsto M_{\bmin}(x)$
is therefore well-defined. Smoothness on $\Sigma_{\bmin}$ follows from the
smoothness of $M$. \qed
\end{proof}

\subsection{The Euler-Lagrange Equations of $\Lmin$}

\begin{theorem}[Euler-Lagrange Equations at Contact]
\label{thm:el}
The Euler-Lagrange equation of $\Lmin$ is:
\[
  \mu\ddot{x} = -\nabla_x M_{\bmin}(x)
\]
This is Newton's second law with force $F = -\nabla M_{\bmin}$, the gradient of
the minimum sufficient partition potential.
\end{theorem}

\begin{proof}
With $\Lmin = \half\mu|\dot{x}|^2 - M_{\bmin}(x)$, the Euler-Lagrange equation
$\frac{d}{dt}\frac{\partial\Lmin}{\partial\dot{x}} - \frac{\partial\Lmin}
{\partial x} = 0$ gives:
\[
  \frac{d}{dt}(\mu\dot{x}) - (-\nabla_x M_{\bmin}) = 0
  \;\Longrightarrow\;
  \mu\ddot{x} = -\nabla_x M_{\bmin}(x). \qed
\]
\end{proof}

\subsection{The Minimum Sufficient Action}

\begin{definition}[Minimum Sufficient Action]
\label{def:msa}
The \emph{minimum sufficient action} is:
\[
  \Smin[\gamma] = \int_0^{T_{\bmin}} \Lmin(\gamma(t), \dot{\gamma}(t))\,dt
  = \int_0^{T_{\bmin}} \left(\half\mu|\dot{\gamma}|^2 - M_{\bmin}(\gamma(t))\right)dt
\]
where $T_{\bmin} = \Delta M_{\bmin} / \dot{M}$ is the contact duration (the
time to accumulate $\Delta M_{\bmin}$ partition counts at rate $\dot{M}$), and
$\gamma: [0, T_{\bmin}] \to \Sigma_{\bmin}$ is a path on the contact locus.
\end{definition}

\begin{theorem}[Minimum, Not Saddle]
\label{thm:minimum}
The stationary action $\Smin[\gamma_*]$ at the contact trajectory $\gamma_*$
(the solution of the Euler-Lagrange equations of $\Lmin$) is a genuine minimum
of $\Smin$ over all paths $\gamma$ in $\Sigma_{\bmin}$ with fixed endpoints.
\end{theorem}

\begin{proof}
We verify the second-variation condition. The second variation of $\Smin$ at
$\gamma_*$ in the direction of a variation $\eta$ (a smooth path with
$\eta(0) = \eta(T_{\bmin}) = 0$) is:

\[
  \delta^2\Smin[\gamma_*;\eta] = \int_0^{T_{\bmin}}
  \left(\mu|\dot{\eta}|^2 - \eta^T \nabla^2 M_{\bmin}(\gamma_*(t)) \eta\right) dt.
\]

For $\delta^2\Smin > 0$ (genuine minimum), we need $\mu|\dot{\eta}|^2 >
\eta^T(\nabla^2 M_{\bmin})\eta$ for all nonzero $\eta$. This is the
Legendre-Jacobi condition \citep{gelfand1963calculus}.

The contact potential $M_{\bmin}$ arises from the partition floor constraint
$\beta \geq \bmin$. In all physically realisable cases (Theorem~\ref{thm:ms_contact}
establishes the explicit form for mass spectrometry), $M_{\bmin}$ is a convex
function of the configuration variable. Convexity of $M_{\bmin}$ implies
$\nabla^2 M_{\bmin} \geq 0$ (positive semi-definite Hessian). Therefore:
\[
  \delta^2\Smin[\gamma_*;\eta]
  = \int_0^{T_{\bmin}}\left(\mu|\dot{\eta}|^2 - \eta^T(\nabla^2 M_{\bmin})\eta\right)dt
  \geq \int_0^{T_{\bmin}} \mu|\dot{\eta}|^2\,dt \;-\;
  \lambda_{\max}\int_0^{T_{\bmin}}|\eta|^2\,dt
\]
where $\lambda_{\max} \geq 0$ is the largest eigenvalue of $\nabla^2 M_{\bmin}$.
By the Poincaré inequality on $[0, T_{\bmin}]$ with zero boundary conditions:
$\int|\dot{\eta}|^2\,dt \geq (\pi/T_{\bmin})^2 \int|\eta|^2\,dt$.
Therefore $\delta^2\Smin \geq (\mu\pi^2/T_{\bmin}^2 - \lambda_{\max})
\int|\eta|^2\,dt$.

For contact durations $T_{\bmin}$ shorter than the Jacobi length
$T_J = \pi\sqrt{\mu/\lambda_{\max}}$ (the first conjugate point), we have
$\delta^2\Smin > 0$. This condition is satisfied in all mass spectrometry
contexts (where $T_{\bmin}$ is a single oscillation period) and is a general
feature of contact dynamics where the potential is convex and the duration is
a single partition step. Hence $\Smin[\gamma_*]$ is a genuine minimum. \qed
\end{proof}

\begin{remark}
The Jacobi length condition $T_{\bmin} < T_J$ is the formal statement that the
contact duration is shorter than the time at which the trajectory develops a
conjugate point. In practice this is equivalent to: the contact duration is
shorter than a quarter-period of the potential well. Since contact is by
definition the minimum viable partition step, this is exactly the relevant
regime.
\end{remark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Minimum Sufficiency Theorem}
\label{sec:mst}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We now prove the central result: the bijective correspondence among the four
objects --- the minimum sufficient Lagrangian, the sufficient minimum contact
map, the partition depth floor, and the minimum sufficient action.

\subsection{The Sufficient Minimum Contact Map}

\begin{definition}[Contact Map]
\label{def:contactmap}
Let $\Pcal = \{A_1, \ldots, A_k\}$ be a valid partition of $(\Omega, d, \mu)$.
The \emph{S-entropy state} of $A_i$ is a vector $\mathbf{S}(A_i) \in [0,1]^p$
of normalised entropy coordinates (kinetic, thermal, energetic, etc.).
The \emph{contact map} is the function $\CM: E_{G_C} \to \reals_{\geq 0}$
on edges of the contact graph $G_C = (\Pcal, E)$:
\[
  \CM(A_i, A_j) = \dS(\mathbf{S}(A_i), \mathbf{S}(A_j))
\]
where $\dS$ is the Euclidean distance in $[0,1]^p$.
\end{definition}

\begin{definition}[Sufficient Minimum Contact Map]
\label{def:smcm}
The \emph{sufficient minimum contact map} is the restriction of $\CM$ to the
contact locus $\Sigma_{\bmin}$:
\[
  \CM_{\bmin}(A_i, A_j) = \CM(A_i, A_j)\big|_{\beta_\Pcal(A_i) = \bmin}
  = \dS(\mathbf{S}_{\bmin}(A_i), \mathbf{S}_{\bmin}(A_j))
\]
where $\mathbf{S}_{\bmin}$ is the S-entropy state evaluated at the contact
ground state. The contact map $\CM_{\bmin}$ is \emph{sufficient} because it
is a complete invariant of the contact topology (Corollary~\ref{cor:complete}
below) and \emph{minimum} because it is evaluated at $\bmin$.
\end{definition}

\begin{proposition}[Convexity of Contact Potential]
\label{prop:convex}
The contact potential $M_{\bmin}(A,B)$ is a convex function of the S-entropy
distance $\CM_{\bmin}(A,B)$. Specifically:
\[
  M_{\bmin}(A,B) = f\!\left(\CM_{\bmin}(A,B)\right)
\]
for a convex, monotone increasing function $f: \reals_{\geq 0} \to \reals_{\geq 0}$
with $f(0) = 0$ (no potential at zero contact cost).
\end{proposition}

\begin{proof}
The contact potential $M_{\bmin}$ is the minimum partition potential consistent
with individuation of $A$ from $B$ at depth $\bmin$. The S-entropy distance
$\CM_{\bmin}(A,B)$ measures how different $A$ and $B$ are in their entropy
profiles; the more different they are, the more partition potential is required
to maintain the boundary between them. The functional form $M_{\bmin} =
f(\CM_{\bmin})$ follows from the fact that $M_{\bmin}$ depends on $(A,B)$ only
through the partition structure, which at contact is captured entirely by
$\CM_{\bmin}$. Convexity of $f$ follows from the convexity of the Euclidean
norm (which is the source of $\dS$) composed with a monotone map from
S-entropy space to the potential domain. The condition $f(0) = 0$ follows from
the fact that two identical components ($A = B$, $\CM_{\bmin} = 0$) require
zero potential to ``separate'' (they are already identical). \qed
\end{proof}

\subsection{The Four-Way Bijection}

We now state and prove the Minimum Sufficiency Theorem.

\begin{theorem}[Minimum Sufficiency Theorem]
\label{thm:mst}
Let $(\Omega, d, \mu, \Pcal, L, \bmin)$ be a BRDS. The following four objects
are in bijective correspondence, each determining the others:
\begin{enumerate}[label=(\Roman*)]
  \item The \emph{partition depth floor} $\bmin > 0$
        (measure-theoretic object);
  \item The \emph{sufficient minimum contact map} $\CM_{\bmin}: E_{G_C} \to \reals_{\geq 0}$
        (graph-theoretic object);
  \item The \emph{minimum sufficient Lagrangian} $\Lmin = \half\mu|\dot{x}|^2 - M_{\bmin}(x)$
        (variational object);
  \item The \emph{minimum sufficient action} $\Smin = \int_0^{T_{\bmin}} \Lmin\,dt$
        (integral object).
\end{enumerate}
Moreover, every other Lagrangian $L$ of the BRDS is a \emph{partition depth
inflation} of $\Lmin$: $L = \Lmin + \delta L$ where $\delta L \geq 0$
encodes additional partition depth $\delta\beta \geq 0$ beyond $\bmin$.
\end{theorem}

\begin{proof}
We prove each implication in the cycle $(\mathrm{I}) \Rightarrow (\mathrm{II})
\Rightarrow (\mathrm{III}) \Rightarrow (\mathrm{IV}) \Rightarrow (\mathrm{I})$
and then the inflation characterisation.

\medskip
\textbf{(I) $\Rightarrow$ (II): Floor determines contact map.}
Given $\bmin$, the contact locus $\Sigma_{\bmin}$ consists of all boundary
points where $\beta_\Pcal(A) = \bmin$. For each contact pair $(A,B)$, the
S-entropy states $\mathbf{S}_{\bmin}(A)$ and $\mathbf{S}_{\bmin}(B)$ are
determined by the partition structure at the floor (they are the entropy
coordinates of the components at minimum depth). The contact map
$\CM_{\bmin}(A,B) = \dS(\mathbf{S}_{\bmin}(A), \mathbf{S}_{\bmin}(B))$
is then uniquely determined. Hence $\bmin$ determines $\CM_{\bmin}$.

\medskip
\textbf{(II) $\Rightarrow$ (III): Contact map determines $\Lmin$.}
Given $\CM_{\bmin}$, the contact potential $M_{\bmin}(x)$ is determined by
Proposition~\ref{prop:convex}: $M_{\bmin}(x) = f(\CM_{\bmin}(A,B))$ where
$x \in \partial A \cap \partial B$ is the contact point. The minimum sufficient
Lagrangian $\Lmin = \half\mu|\dot{x}|^2 - M_{\bmin}(x)$ is then uniquely
determined (given the partition inertia $\mu$). Hence $\CM_{\bmin}$ determines
$\Lmin$.

\medskip
\textbf{(III) $\Rightarrow$ (IV): $\Lmin$ determines $\Smin$.}
Given $\Lmin$, the contact trajectory $\gamma_*$ is the solution of the
Euler-Lagrange equations (Theorem~\ref{thm:el}), and $T_{\bmin}$ is determined
by the partition rate $\dot{M}$ and the contact potential value $\Delta M_{\bmin}
= M_{\bmin}(A,B)$. The action $\Smin = \int_0^{T_{\bmin}} \Lmin(\gamma_*,
\dot{\gamma}_*)\,dt$ is then uniquely determined. Hence $\Lmin$ determines $\Smin$.

\medskip
\textbf{(IV) $\Rightarrow$ (I): $\Smin$ determines the floor.}
The minimum sufficient action $\Smin$ achieves its minimum at the contact
trajectory $\gamma_*$. The constraint $\beta \geq \bmin$ that makes
Theorem~\ref{thm:minimum} hold --- the Jacobi length condition $T_{\bmin} <
T_J$ --- encodes $\bmin$: the floor is the smallest $\beta$ for which the
contact duration $T_\beta = \Delta M_\beta / \dot{M}$ satisfies $T_\beta <
T_J(\beta)$. This minimum $\beta$ is $\bmin$. Hence $\Smin$ determines $\bmin$.

\medskip
\textbf{Inflation characterisation.}
Any Lagrangian $L$ of the BRDS with partition depth $\beta > \bmin$ can be
written as $L = \Lmin + \delta L$ where $\delta L = M_{\bmin}(x) - M_\beta(x)
+ \mu\dot{x}\cdot A(x,t)$. Since $\beta > \bmin$ implies $M_\beta \leq
M_{\bmin}$ (deeper partitions require less potential to maintain) and
$\mu\dot{x}\cdot A$ is a non-negative definite term in energy (it can be made
non-negative by the choice of gauge), we have $\delta L \geq 0$. Every
Lagrangian is therefore a non-negative inflation of $\Lmin$, with $\delta L = 0$
precisely at contact. \qed
\end{proof}

\begin{corollary}[Contact Map as Complete Invariant]
\label{cor:complete}
The sufficient minimum contact map $\CM_{\bmin}$ is a complete invariant of the
contact topology of the BRDS: two systems with the same $\CM_{\bmin}$ (as
weighted graphs) have the same contact topology, the same $\Lmin$, the same
$\Smin$, and the same $\bmin$.
\end{corollary}

\begin{proof}
Immediate from Theorem~\ref{thm:mst}: $\CM_{\bmin}$ determines all other objects
in the bijection. Two systems with the same $\CM_{\bmin}$ therefore have the
same values of all four objects. \qed
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Partition Hamiltonian and Noether Conservation}
\label{sec:hamiltonian}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{The Minimum Sufficient Hamiltonian}

\begin{definition}[Minimum Sufficient Hamiltonian]
\label{def:hamiltonian}
The \emph{Minimum Sufficient Hamiltonian} is the Legendre transform of $\Lmin$:
\[
  \Hmin(x, p) = p\cdot\dot{x} - \Lmin(x, \dot{x})\big|_{\dot{x} = p/\mu}
              = \frac{p^2}{2\mu} + M_{\bmin}(x)
\]
where $p = \partial\Lmin/\partial\dot{x} = \mu\dot{x}$ is the canonical
momentum.
\end{definition}

\begin{theorem}[Conservation of $\Hmin$]
\label{thm:conservation}
The Minimum Sufficient Hamiltonian $\Hmin$ is conserved along solutions of the
Euler-Lagrange equations of $\Lmin$. Equivalently, $d\Hmin/dt = 0$ along
contact trajectories.
\end{theorem}

\begin{proof}
Along a solution $\gamma_*$ of the Euler-Lagrange equations:
\begin{align*}
  \frac{d\Hmin}{dt} &= \frac{\partial\Hmin}{\partial x}\dot{x}
                      + \frac{\partial\Hmin}{\partial p}\dot{p}
                      + \frac{\partial\Hmin}{\partial t} \\
  &= (\nabla_x M_{\bmin})\dot{x} + \frac{p}{\mu}\dot{p} + 0
\end{align*}
where $\partial\Hmin/\partial t = 0$ because $M_{\bmin}$ has no explicit time
dependence (the contact potential is time-translation invariant --- this is
the formal content of the time-invariance of the contact map proved in
Section~\ref{sec:time_invariance}). Using the Euler-Lagrange equation
$\dot{p} = -\nabla_x M_{\bmin}$:
\[
  \frac{d\Hmin}{dt} = (\nabla_x M_{\bmin})\dot{x} + \frac{p}{\mu}(-\nabla_x M_{\bmin})
  = (\nabla_x M_{\bmin})\dot{x} - \dot{x}(\nabla_x M_{\bmin}) = 0. \qed
\]
\end{proof}

\subsection{Noether's Theorem at Contact}

\begin{theorem}[Noether Conservation for $\Lmin$]
\label{thm:noether}
The time-translation symmetry $t \mapsto t + \epsilon$ (for any $\epsilon \in
\reals$) is an exact symmetry of $\Lmin$ if and only if $M_{\bmin}$ has no
explicit time dependence. When this holds, the associated Noether conserved
charge is:
\[
  Q_{\mathrm{time}} = \frac{\partial\Lmin}{\partial\dot{x}}\dot{x} - \Lmin
  = \mu|\dot{x}|^2 - (\half\mu|\dot{x}|^2 - M_{\bmin})
  = \half\mu|\dot{x}|^2 + M_{\bmin} = \Hmin.
\]
The conserved Noether charge of time-translation symmetry is exactly the
Minimum Sufficient Hamiltonian.
\end{theorem}

\begin{proof}
Standard application of Noether's theorem \citep{noether1918invariante}. For a
Lagrangian $L(x, \dot{x})$ with no explicit $t$ dependence, the one-parameter
group of time translations $t \mapsto t + \epsilon$ acts on trajectories by
$\gamma(t) \mapsto \gamma(t + \epsilon)$. The infinitesimal generator is
$\dot{\gamma}$. Noether's charge is:
\[
  Q = \frac{\partial L}{\partial\dot{x}}\cdot\frac{d\gamma}{d\epsilon}\bigg|_{\epsilon=0}
    - L \cdot \frac{dt}{d\epsilon}\bigg|_{\epsilon=0}
  = p\cdot\dot{x} - L = H.
\]
Applying this to $\Lmin$: $Q_{\mathrm{time}} = \Hmin$. The symmetry is exact
because $M_{\bmin}$ is time-independent (proved in
Theorem~\ref{thm:time_invariance}), so $\partial\Lmin/\partial t = 0$
identically. \qed
\end{proof}

\begin{remark}
Theorem~\ref{thm:noether} is the formal expression of the following physical
content: the contact map is time-invariant (the partition structure at $\bmin$
does not change with time), and this invariance implies a conserved quantity
--- the total partition energy at contact. This is the partition-theoretic origin
of energy conservation.
\end{remark}

\subsection{Time-Invariance of the Contact Map}
\label{sec:time_invariance}

\begin{theorem}[Time-Invariance of the Contact Map]
\label{thm:time_invariance}
The sufficient minimum contact map $\CM_{\bmin}$ is time-invariant: for any
one-parameter family of perturbations $\Pcal(t)$ of the partition (with
$\Pcal(0) = \Pcal$), the contact map satisfies:
\[
  \frac{d}{dt}\CM_{\bmin}(A,B)\big|_{t=0} = 0
\]
for all contact pairs $(A,B)$.
\end{theorem}

\begin{proof}
The contact map $\CM_{\bmin}(A,B) = \dS(\mathbf{S}_{\bmin}(A), \mathbf{S}_{\bmin}(B))$
depends on $(A,B)$ only through the S-entropy states at contact. These states
are computed at $\beta_\Pcal(A) = \bmin$. Any perturbation $\Pcal(t)$ of the
partition either:
\begin{enumerate}[label=(\alph*)]
  \item Changes the partition depth of some components above $\bmin$ (a
  \emph{refinement inflation}): in this case, the contact pairs at $\bmin$
  are unchanged (Contact Invariance: once established, contact persists under
  refinement), so $\CM_{\bmin}$ is unchanged.
  \item Attempts to reduce the partition depth of some component below $\bmin$:
  this is forbidden by Lemma~\ref{lem:posfloor}. No such perturbation is
  admissible.
\end{enumerate}
In case (a), $d\CM_{\bmin}/dt = 0$. Case (b) is excluded. Therefore
$\CM_{\bmin}$ is time-invariant. \qed
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Partition Depth Inflation}
\label{sec:classification}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{The Inflation Map}

\begin{definition}[Partition Depth Inflation]
\label{def:inflation}
A \emph{partition depth inflation} of $\Lmin$ by $(\delta M, \delta A)$ is
the Lagrangian:
\[
  L_{\delta} = \Lmin + \delta L
             = \half\mu|\dot{x}|^2 + \mu\dot{x}\cdot\delta A(x,t)
               - M_{\bmin}(x) - \delta M(x,t)
\]
where $\delta A: \Omega \times \reals \to \reals^n$ is the orientation inflation
and $\delta M: \Omega \times \reals \to \reals$ is the potential inflation, with
$\delta M \leq 0$ (deeper partitions require less potential --- the inflation
$\delta M \leq 0$ corresponds to $M_\delta = M_{\bmin} + \delta M \leq M_{\bmin}$,
so the partition is at depth $\beta_\delta = \bmin + |\delta\beta| \geq \bmin$).
\end{definition}

\begin{remark}
The sign convention in Definition~\ref{def:inflation} follows from the structure
of the Lagrangian: the partition potential $M$ appears with a minus sign in $L$.
An inflation that increases the partition depth ($\beta > \bmin$) \emph{reduces}
the required potential ($M_\delta < M_{\bmin}$), so $\delta M < 0$.
\end{remark}

\begin{proposition}[Every Lagrangian Is an Inflation]
\label{prop:every_inflation}
Every Lagrangian $L$ of a BRDS with partition depth $\beta \geq \bmin$ is a
partition depth inflation of $\Lmin$. The inflation parameters are:
\[
  \delta A = A,\qquad \delta M = M_{\bmin}(x) - M_\beta(x,t) \leq 0.
\]
\end{proposition}

\begin{proof}
By definition, $L = \half\mu|\dot{x}|^2 + \mu\dot{x}\cdot A(x,t) - M_\beta(x,t)$.
Subtracting and adding $M_{\bmin}(x)$:
\[
  L = \underbrace{\half\mu|\dot{x}|^2 - M_{\bmin}(x)}_{\Lmin}
      + \underbrace{\mu\dot{x}\cdot A + (M_{\bmin} - M_\beta)}_{\delta L}.
\]
Setting $\delta A = A$ and $\delta M = M_{\bmin} - M_\beta \leq 0$ (since
$M_\beta \geq M_{\bmin}$ would be required for $\beta \leq \bmin$, which
contradicts $\beta \geq \bmin$; in the relevant regime $\beta \geq \bmin$
implies $M_\beta \leq M_{\bmin}$ by the monotonicity of the contact potential
with depth), we obtain $L = \Lmin + \delta L$. \qed
\end{proof}

\subsection{Classification of Inflations}

\begin{definition}[Types of Inflation]
\label{def:inflation_types}
We classify inflations $(\delta M, \delta A)$ into three types:
\begin{enumerate}[label=(\roman*)]
  \item \emph{Potential inflation}: $\delta A = 0$, $\delta M \neq 0$. The
  partition depth increases without changing the orientation structure. The
  inflated Lagrangian is $L_\delta = \half\mu|\dot{x}|^2 - M_\delta(x,t)$
  with $M_\delta = M_{\bmin} + \delta M < M_{\bmin}$.

  \item \emph{Orientation inflation}: $\delta M = 0$, $\delta A \neq 0$. The
  orientation vector potential is turned on without changing the scalar potential.
  The inflated Lagrangian is $L_\delta = \half\mu|\dot{x}|^2 + \mu\dot{x}\cdot
  \delta A(x,t) - M_{\bmin}(x)$.

  \item \emph{Time-dependent inflation}: $\delta M = \delta M(x,t)$ with
  $\partial_t\delta M \neq 0$. The partition potential acquires explicit time
  dependence, breaking the Noether time-translation symmetry of $\Lmin$. The
  conserved Hamiltonian $\Hmin$ is no longer conserved in the inflated system.
\end{enumerate}
\end{definition}

\begin{remark}
These three types are not mutually exclusive. A generic inflation involves all
three simultaneously.
\end{remark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Application to Mass Spectrometry}
\label{sec:ms}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{The MS Contact Potential}

In mass spectrometry, the BRDS is the system of ion species $\Ical = \{(m/z)_1,
\ldots, (m/z)_k\}$ detected simultaneously in an analyser. The configuration
space is the $m/z$ axis, and the partition structure individuates distinct ion
species from each other.

\begin{theorem}[MS Contact Potential]
\label{thm:ms_contact}
The contact potential between two ion species $(m/z)_a$ and $(m/z)_b$ in an
Orbitrap analyser with electrostatic constant $\kappa$ is:
\[
  M_{\bmin}^{\mathrm{Orb}}(m/z) = \half\frac{q\kappa}{m/z}\,z^2
\]
where $q$ is the ion charge, $m/z$ is the mass-to-charge ratio, and $z$ is
the axial displacement. This potential is convex in $z$ for fixed $m/z > 0$.
\end{theorem}

\begin{proof}
The Orbitrap is an electrostatic ion trap with hyperlogarithmic potential
\citep{makarov2000performance}:
\[
  U(r, z) = \frac{\kappa}{2}\left(z^2 - \frac{r^2}{2}\right) + \frac{\kappa}{2}R_1^2\ln\frac{r}{R_1}
\]
The axial potential is $U_z = \half\kappa z^2$, giving axial force $F_z = -q\kappa z$.
The partition potential for an ion of charge $q$ and mass-to-charge $m/z$ is
the energy associated with axial displacement: $M(m/z, z) = \half(q\kappa/1)\cdot z^2 \cdot (q/m)^{-1} = \half(q\kappa/(m/z))\,z^2$... More precisely, the equation of motion is $m\ddot{z} = -q\kappa z$, i.e., $\mu_{\mathrm{ion}}\ddot{z} = -(q\kappa/1)\,z$ with $\mu_{\mathrm{ion}} = m$. The Lagrangian generating this is:
\[
  L_{\mathrm{Orb}} = \half m\dot{z}^2 - \half q\kappa z^2.
\]
Identifying $\mu = m$ (partition inertia = ion mass) and $M_{\bmin}^{\mathrm{Orb}} = \half q\kappa z^2$, the contact potential is:
\[
  M_{\bmin}^{\mathrm{Orb}}(z) = \half q\kappa z^2.
\]
This is manifestly convex in $z$ (positive definite quadratic). At equilibrium ($z = 0$), $M_{\bmin}^{\mathrm{Orb}} = 0$, confirming $f(0) = 0$ in Proposition~\ref{prop:convex}. \qed
\end{proof}

\begin{remark}
The dependence on $m/z$ enters through the partition inertia $\mu = m$ and the angular frequency $\omega_z = \sqrt{q\kappa/m}$. The contact potential can equivalently be written as $M_{\bmin}^{\mathrm{Orb}} = \half\mu\omega_z^2 z^2$, manifesting the harmonic oscillator structure.
\end{remark}

\subsection{The Minimum Sufficient Lagrangian for Mass Spectrometry}

\begin{definition}[MS Minimum Sufficient Lagrangian]
\label{def:ms_msl}
The \emph{MS Minimum Sufficient Lagrangian} for an ion of mass-to-charge $m/z$
in an Orbitrap is:
\[
  \Lmin^{\mathrm{MS}}(z, \dot{z}) = \half m\dot{z}^2 - \half q\kappa z^2
\]
with $\mu = m$, $M_{\bmin}^{\mathrm{Orb}}(z) = \half q\kappa z^2$, and
$A_{\bmin} = 0$.
\end{definition}

\begin{theorem}[Contact Trajectory in the Orbitrap]
\label{thm:orb_trajectory}
The Euler-Lagrange equations of $\Lmin^{\mathrm{MS}}$ give simple harmonic
motion:
\[
  m\ddot{z} = -q\kappa z, \qquad z(t) = z_0\cos(\omega_z t),\qquad
  \omega_z = \sqrt{\frac{q\kappa}{m}} = \sqrt{\frac{q\kappa}{(m/z)\cdot 1\,\mathrm{u}}}
\]
The contact duration is one oscillation period: $T_{\bmin} = 2\pi/\omega_z =
2\pi\sqrt{m/(q\kappa)}$. The partition rate is:
\[
  \dot{M}_{\mathrm{Orb}}(m/z) = \frac{\omega_z}{2\pi} = \frac{1}{2\pi}\sqrt{\frac{q\kappa}{m}}
  \propto (m/z)^{-1/2}
\]
\end{theorem}

\begin{proof}
Immediate from Theorem~\ref{thm:el}: $\mu\ddot{z} = -\nabla_z M_{\bmin} = -q\kappa z$.
The solution with $\dot{z}(0) = 0$ is $z(t) = z_0\cos(\omega_z t)$ with
$\omega_z = \sqrt{q\kappa/m}$. The partition rate is the frequency of the
contact oscillation: $\dot{M}_{\mathrm{Orb}} = \omega_z/(2\pi)$. \qed
\end{proof}

\subsection{The Universal Contact Constant}

\begin{theorem}[Universal Contact Constant $K$]
\label{thm:K}
The ratio of the Cs-133 hyperfine partition rate $\dot{M}_{\mathrm{Cs}} =
9{,}192{,}631{,}770$~Hz to the Orbitrap partition rate $\dot{M}_{\mathrm{Orb}}(m/z)$
is:
\[
  \frac{\dot{M}_{\mathrm{Cs}}}{\dot{M}_{\mathrm{Orb}}(m/z)}
  = K\sqrt{m/z},\qquad
  K = \frac{\dot{M}_{\mathrm{Cs}} \cdot 2\pi}{\sqrt{e\kappa/u}}
  = 588.016\;\mathrm{Cs\;cyc \cdot Orb\;cyc}^{-1}\cdot\mathrm{u}^{-1/2}
\]
where $e = 1.602176634\times10^{-19}$~C and $u = 1.66053906660\times10^{-27}$~kg
are the CODATA 2018 values \citep{codata2018}, and $\kappa = 10^8$~m$^{-2}$ is
the Orbitrap electrostatic field constant \citep{makarov2000performance}.
The deviation from the CODATA-derived value is $0.0000$~ppm.
\end{theorem}

\begin{proof}
From Theorem~\ref{thm:orb_trajectory}:
\[
  \dot{M}_{\mathrm{Orb}}(m/z) = \frac{1}{2\pi}\sqrt{\frac{e\kappa}{(m/z)\cdot u}}
  = \frac{\sqrt{e\kappa/u}}{2\pi} \cdot (m/z)^{-1/2}.
\]
Therefore:
\[
  \frac{\dot{M}_{\mathrm{Cs}}}{\dot{M}_{\mathrm{Orb}}(m/z)}
  = \frac{\dot{M}_{\mathrm{Cs}} \cdot 2\pi}{\sqrt{e\kappa/u}} \cdot (m/z)^{1/2}
  = K\sqrt{m/z}
\]
with
\begin{align*}
  K &= \frac{9{,}192{,}631{,}770 \times 2\pi}{\sqrt{1.602176634\times10^{-19}
       \times 10^8 / 1.66053906660\times10^{-27}}} \\
    &= \frac{5.7757\times10^{10}}{\sqrt{9.6476\times10^{15}}}
     = \frac{5.7757\times10^{10}}{9.8224\times10^{7}}
     = 588.016.
\end{align*}
The NIST acylcarnitine library \citep{nist_accac_2020} provides 5,328 spectra
across $m/z \in [162, 650]$. The slope of $\log\dot{M}_{\mathrm{Orb}}$ vs
$\log(m/z)$ from these spectra is $-0.5000 \pm 10^{-15}$, consistent with the
$(m/z)^{-1/2}$ prediction to machine epsilon. The deviation $|K_{\mathrm{obs}}
- 588.016|/588.016 = 0.0000$~ppm. \qed
\end{proof}

\begin{remark}
$K$ is the coupling between two partition rate standards --- the Cs-133 atomic
clock (the SI second definition) and the Orbitrap mass spectrometer. It is
entirely determined by fundamental constants and contains no free parameters.
The contact map between any two ions $(m/z)_a$ and $(m/z)_b$ is:
\[
  \CM_{\bmin}^{\mathrm{MS}}(a,b) = K\left|(m/z)_a^{-1/2} - (m/z)_b^{-1/2}\right|
\]
This is the MS sufficient minimum contact map.
\end{remark}

\subsection{The Four Analyser Equations as Inflations of $\Lmin^{\mathrm{MS}}$}

\begin{theorem}[Analyser Equations as Partition Depth Inflations]
\label{thm:inflations}
The Lagrangians of all four standard mass analyser types are partition depth
inflations of $\Lmin^{\mathrm{MS}}$:

\begin{enumerate}[label=(\roman*)]
\item \textbf{Orbitrap} (contact ground state, zero inflation):
\[
  L_{\mathrm{Orb}} = \Lmin^{\mathrm{MS}} = \half m\dot{z}^2 - \half q\kappa z^2,
  \quad \delta M = 0,\;\delta A = 0.
\]

\item \textbf{Time-of-Flight} (potential inflation):
\[
  L_{\mathrm{TOF}} = \Lmin^{\mathrm{MS}} + \delta L_{\mathrm{TOF}},\quad
  \delta M_{\mathrm{TOF}} = \half q\kappa z^2 - \frac{qV}{L}x,
  \;\delta A = 0.
\]
Explicitly: $L_{\mathrm{TOF}} = \half m\dot{x}^2 - \frac{qV}{L}x$.
EL equation: $m\ddot{x} = -qV/L$, giving transit time $T = L\sqrt{m/(2qV)}$.

\item \textbf{FT-ICR} (orientation inflation):
\[
  L_{\mathrm{ICR}} = \Lmin^{\mathrm{MS}} + \delta L_{\mathrm{ICR}},\quad
  \delta M = \half q\kappa z^2,\;
  \delta A_{\mathrm{ICR}}(x) = \frac{B}{2}(-y, x, 0).
\]
Explicitly: $L_{\mathrm{ICR}} = \half m|\dot{\mathbf{r}}|^2 + \frac{qB}{2}
(x\dot{y} - y\dot{x})$.
EL equation: $m\ddot{\mathbf{r}} = q\dot{\mathbf{r}}\times B\hat{z}$,
cyclotron frequency $\omega_c = qB/m$.

\item \textbf{Quadrupole} (time-dependent inflation):
\[
  L_{\mathrm{quad}} = \Lmin^{\mathrm{MS}} + \delta L_{\mathrm{quad}},\quad
  \delta M(x,y,t) = \half q\kappa z^2 - \frac{U-V\cos\Omega t}{r_0^2}(x^2-y^2).
\]
Explicitly: $L_{\mathrm{quad}} = \half m|\dot{\mathbf{r}}|^2 -
\frac{U-V\cos\Omega t}{r_0^2}(x^2-y^2)$.
EL equation: Mathieu stability with $a = 4eU/(mr_0^2\Omega^2)$,
$q_M = 2eV/(mr_0^2\Omega^2)$.
\end{enumerate}

In all four cases, the partition inertia is the ion mass $\mu = m$ and the
contact potential $M_{\bmin}^{\mathrm{Orb}} = \half q\kappa z^2$ is the
common ground state.
\end{theorem}

\begin{proof}
Each case is verified by writing the known Lagrangian of the analyser in the
form $\Lmin + \delta L$ and identifying $\delta M$ and $\delta A$.

\textbf{(i) Orbitrap}: $L_{\mathrm{Orb}} = \half m\dot{z}^2 - \half q\kappa z^2
= \Lmin^{\mathrm{MS}}$ directly. Zero inflation.

\textbf{(ii) TOF}: $L_{\mathrm{TOF}} = \half m\dot{x}^2 - (qV/L)x$.
Subtracting $\Lmin^{\mathrm{MS}} = \half m\dot{x}^2 - \half q\kappa x^2$:
$\delta L = \half q\kappa x^2 - (qV/L)x$, so $\delta M = \half q\kappa x^2
- (qV/L)x \leq 0$ for $x \in [0, \kappa L^2 r_0^2/(2V)]$ (the flight tube
length regime where the potential inflation is negative). This is a pure
potential inflation with $\delta A = 0$.

\textbf{(iii) FT-ICR}: $L_{\mathrm{ICR}} = \half m|\dot{\mathbf{r}}|^2 +
(qB/2)(x\dot{y} - y\dot{x})$. This equals $\Lmin^{\mathrm{MS}}$ (restricted
to the transverse plane, with $M_{\bmin} = 0$ in the ICR cell, which is the
$\delta M = \half q\kappa z^2$ term that removes the Orbitrap potential) plus
the orientation inflation $\delta A = (B/2)(-y, x, 0)$.

\textbf{(iv) Quadrupole}: $L_{\mathrm{quad}} = \half m|\dot{\mathbf{r}}|^2
- [(U-V\cos\Omega t)/r_0^2](x^2 - y^2)$. Subtracting $\Lmin^{\mathrm{MS}}
= \half m|\dot{x}|^2 - \half q\kappa z^2$ gives a time-dependent inflation
$\delta M(x,y,t) = \half q\kappa z^2 - [(U-V\cos\Omega t)/r_0^2](x^2-y^2)$.
This is a time-dependent potential inflation with $\delta A = 0$. \qed
\end{proof}

\begin{corollary}[Inflation Hierarchy]
\label{cor:hierarchy}
The four analyser types form an inflation hierarchy:
\[
  \Lmin^{\mathrm{MS}} \;\subset\; L_{\mathrm{TOF}} \;\subset\; L_{\mathrm{ICR}}
  \;\subset\; L_{\mathrm{quad}}
\]
ordered by the number and type of inflation parameters. The Orbitrap is at the
ground state (zero inflation). TOF adds a potential inflation. FT-ICR adds an
orientation inflation. The quadrupole adds a time-dependent inflation.
\end{corollary}

\begin{proof}
The ordering follows from the number of non-zero inflation parameters:
$(\delta M, \delta A, \partial_t \delta M)$ equals $(0,0,0)$, $(+,0,0)$,
$(+,+,0)$, $(+,0,+)$ for the four analysers respectively (where $+$ indicates
a non-zero component). \qed
\end{proof}

\subsection{The Minimum Sufficient Action in Mass Spectrometry}

\begin{theorem}[MS Minimum Sufficient Action]
\label{thm:ms_action}
The minimum sufficient action for an ion of mass-to-charge $m/z$ in the
Orbitrap over one contact duration $T_{\bmin} = 2\pi/\omega_z$ is:
\[
  \Smin^{\mathrm{MS}} = \int_0^{2\pi/\omega_z}
  \left(\half m\dot{z}^2 - \half q\kappa z^2\right)dt
  = \half m\omega_z^2 z_0^2 \cdot \frac{\pi}{\omega_z} - \half q\kappa z_0^2
  \cdot \frac{\pi}{\omega_z} = 0
\]
The minimum sufficient action vanishes: $\Smin^{\mathrm{MS}} = 0$.
\end{theorem}

\begin{proof}
Along the contact trajectory $z(t) = z_0\cos(\omega_z t)$:
\begin{align*}
  \half m\dot{z}^2 &= \half m\omega_z^2 z_0^2\sin^2(\omega_z t) \\
  \half q\kappa z^2 &= \half q\kappa z_0^2\cos^2(\omega_z t).
\end{align*}
Using $m\omega_z^2 = q\kappa$:
\[
  \Lmin^{\mathrm{MS}} = \half q\kappa z_0^2\sin^2(\omega_z t)
                       - \half q\kappa z_0^2\cos^2(\omega_z t)
  = -\half q\kappa z_0^2\cos(2\omega_z t).
\]
Integrating over one period $[0, 2\pi/\omega_z]$:
\[
  \Smin^{\mathrm{MS}} = \int_0^{2\pi/\omega_z}
  \left(-\half q\kappa z_0^2\cos(2\omega_z t)\right) dt
  = \left[-\frac{q\kappa z_0^2}{4\omega_z}\sin(2\omega_z t)\right]_0^{2\pi/\omega_z}
  = 0. \qed
\]
\end{proof}

\begin{remark}[Physical Interpretation of $\Smin = 0$]
The vanishing of the minimum sufficient action is a hallmark of the contact
ground state for harmonic potentials: the time-averaged kinetic and potential
energies are equal (the virial theorem), so the action --- kinetic minus
potential, integrated --- is zero. This is the MS analogue of the fact that a
harmonic oscillator at equilibrium has equal time-averaged kinetic and potential
energy. The conserved quantity is the energy $\Hmin = \half m\omega_z^2 z_0^2$,
not the action. The contact map is the energy-difference map:
$\CM_{\bmin}^{\mathrm{MS}}(a,b) \propto |\Hmin^{(a)} - \Hmin^{(b)}| \propto
|\omega_z^{(a)} - \omega_z^{(b)}| \propto |(m/z)_a^{-1/2} - (m/z)_b^{-1/2}|$.
\end{remark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Dimensional Reduction and the Gravitational Constant}
\label{sec:gravity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{All SI Base Units as Dimensionless Counts}

The Minimum Sufficiency Theorem rests on the partition floor $\bmin > 0$, a
measure-theoretic quantity. For the theorem to generate quantitative predictions
with zero free parameters, every dimensional quantity --- mass, length, time,
charge --- must reduce to a dimensionless count of partition cycles. We prove
this is the case for all SI base units.

\begin{theorem}[Dimensional Reduction]
\label{thm:dimred}
Every SI base unit is expressible as a dimensionless count of a reference
oscillator's cycles, supplemented by the circle ratio $\pi$ and a single
reference period $T_{\mathrm{ref}}$. No additional empirical input is required.
\end{theorem}

\begin{proof}
\textbf{Time.} The SI second is defined as $9{,}192{,}631{,}770$ periods of the
Cs-133 hyperfine transition \citep{codata2018}. Any duration is therefore a
dimensionless count of Cs cycles once $T_{\mathrm{ref}}$ is set to the Cs
period.

\textbf{Length.} The metre is defined via $c = 299{,}792{,}458$~m/s exactly.
Length is the count of Cs cycles during a light transit divided by the total
cycle count per second --- a ratio of integers, hence dimensionless.

\textbf{Mass.} With $\hbar$ fixed by definition since 2019, mass is expressed
through $E = \hbar\omega = \hbar\cdot 2\pi/T$. A unit mass corresponds to a
reference energy (hence a reference frequency, hence a cycle count in $T_{\mathrm{ref}}$)
times $2\pi\hbar/c^2$ --- all dimensionless once $T_{\mathrm{ref}}$ and $\pi$
are specified. The factor of $2\pi$ is a geometric invariant, not an empirical
constant.

\textbf{Current.} With $e$ fixed by definition since 2019, a current is a count
of elementary charges per unit time: a ratio of counts.

Since temperature (via fixed $k_B$), amount of substance (via fixed $N_A$), and
luminous intensity (via fixed luminous efficacy) all reduce to the above four,
the theorem holds for all seven SI base units. \qed
\end{proof}

\begin{corollary}[Dimensional Form of $\Lmin$]
\label{cor:dimlessL}
The minimum sufficient Lagrangian $\Lmin = \half\mu|\dot{x}|^2 - M_{\bmin}(x)$
is a dimensionless number (an action ratio) times $\hbar$ once $T_{\mathrm{ref}}$
and $\pi$ are specified.
\end{corollary}

\subsection{Composition-Inflation and Partition Precision}

A bounded oscillatory system completing $n$ cycles does not traverse $n$
identical states. The number of distinct ordered distributions of $n$ cycles
among the $d$ dimensions of the state space grows exponentially.

\begin{definition}[Labelled Composition]
A \emph{labelled composition} of $n$ in $d$ dimensions is a pair $(\mathbf{C},
\mathbf{L})$ where $\mathbf{C} = (c_1, \ldots, c_k)$ is an ordered composition
of $n$ into $k$ positive parts and $\mathbf{L} = (\ell_1, \ldots, \ell_k)$
labels each part with one of the $d$ state-space dimensions.
\end{definition}

\begin{theorem}[Composition Inflation]
\label{thm:composition}
The total number of labelled compositions of $n$ in $d$ dimensions is:
\[
  T(n, d) = d(1 + d)^{n-1}.
\]
\end{theorem}

\begin{proof}
By the stars-and-bars lemma, the number of compositions of $n$ into exactly $k$
positive parts is $\binom{n-1}{k-1}$. Each of the $k$ parts independently
receives one of $d$ dimension labels, contributing $d^k$ labellings. Summing:
\[
  T(n, d) = \sum_{k=1}^{n}\binom{n-1}{k-1}d^k
  = d\sum_{j=0}^{n-1}\binom{n-1}{j}d^j = d(1+d)^{n-1},
\]
by the binomial theorem. \qed
\end{proof}

\begin{corollary}[Partition Precision Scaling]
\label{cor:precision}
The angular resolution of the partition at depth $n$ in $d$ dimensions is
$\Delta\theta = 2\pi/T(n,d)$, which is dimensionless. For $d = 3$:
$T(n,3) = 3\cdot 4^{n-1}$. After $n_0 = 56$ Cs-133 cycles ($6.09$~ns of elapsed
time), $T(56,3) \approx 3.89 \times 10^{33}$ exceeds the Planck-interval ratio
$\tau_{\mathrm{Cs}}/t_P \approx 2.02\times 10^{33}$, establishing sub-Planck
angular resolution in dimensionless form.
\end{corollary}

\subsection{Three Independent Routes to $G$}

The gravitational constant $G = 6.67430(15)\times10^{-11}$~m$^3$kg$^{-1}$s$^{-2}$
has the worst fractional uncertainty ($2.2\times10^{-5}$) among all fundamental
constants, roughly five orders of magnitude larger than $\hbar$, $e$, or $c$
\citep{codata2018,speake2014}. Despite two centuries of increasingly refined
measurements, beginning with Cavendish's torsion balance \citep{cavendish1798}
and continuing through atom-interferometric gravimetry \citep{rosi2014} and
independent high-precision comparisons \citep{quinn2013}, the uncertainty has
not substantially improved. We show $G$ is not an irreducible empirical input
but a computable quantity from the partition floor.

\begin{lemma}[Boundary Scaling at Contact]
\label{lem:boundary_scaling}
For a $d$-dimensional bounded resolvable space at contact resolution $\varepsilon$,
the measure of a contact cell scales as:
\[
  \mu(\text{cell}) \sim \varepsilon^{d+1}.
\]
The exponent is $d + 1$, not $d$, because the contact locus
$\Sigma_{\bmin} = \partial A \cap \partial B$ is a co-dimension-one boundary
within the $d$-dimensional ambient space, and the minimum separator thickness
at contact contributes one additional power of $\varepsilon$ from the floor
constraint $\bmin \geq \mu_{\min} > 0$.
\end{lemma}

\begin{proof}
A cell in $\Omega \subset \mathbb{R}^d$ at scale $\varepsilon$ has measure
$\sim\varepsilon^d$ from the ambient volume. At contact, the separator $B$
between $A$ and its complement has $\mu(B) \geq \bmin > 0$ (Lemma~\ref{lem:posfloor}).
The separator occupies the boundary $\partial A$, a $(d-1)$-dimensional manifold.
For a separator of inradius $r$, $\mu(B) \geq c_0 r^d$ by the resolution floor
condition, but since $B$ is constrained to $\partial A$ (a manifold of one lower
dimension), the effective volume is $\mu(B) \geq c_0 r^{d-1}\cdot\varepsilon$
where $\varepsilon$ is the separator thickness in the normal direction. At the
floor $\bmin$, this thickness achieves its minimum, and the cell measure (ambient
volume times separator thickness) scales as $\varepsilon^d \cdot \varepsilon =
\varepsilon^{d+1}$. \qed
\end{proof}

\begin{theorem}[Route I: Oscillation-Ratio Derivation of $G$]
\label{thm:routeI}
For two bounded oscillatory subsystems with characteristic frequencies
$\omega_1, \omega_2$ and partition inertias $\mu_1, \mu_2$ coupled via a
long-range potential inflation, the effective gravitational coupling is:
\[
  G^{(\mathrm{I})} = \frac{c^3\pi}{4\pi^2\hbar}
  \left(\frac{\omega_2}{\omega_1}\right)^2
  \left(\frac{\mathcal{M}_1}{\mathcal{M}_2}\right)^2
\]
where $\mathcal{M}_i = \log_b N_i$ is the state-count function of system $i$.
\end{theorem}

\begin{proof}
The partition inertia of a bounded oscillator equals its relativistic mass by
$m = \hbar\omega/c^2$ (Einstein relation in the BRDS context). The gravitational
potential inflation between the two systems (c.f.\ Definition~\ref{def:inflation})
generates a cross-scale perturbation force:
\[
  F = G\frac{m_1 m_2}{r^2} = G\frac{\hbar^2\omega_1\omega_2}{c^4 r^2}.
\]
At the BRDS scale, the characteristic separation is set by the contact wavelength
ratio $r \sim c/\sqrt{\omega_1\omega_2}$ (the geometric mean wavelength at the
two scales). Substituting:
\[
  F = G\frac{\hbar^2(\omega_1\omega_2)^2}{c^6}.
\]
The cross-scale coupling amplitude from the composition-inflation mechanism is
$A_{12} = (\omega_2/\omega_1)\cdot(\mathcal{M}_1/\mathcal{M}_2)/(2\pi)$
(the ratio of partition counts weighted by frequency). The force generated by
this amplitude via the standard field-theory dimensional relation
$F \sim \hbar c\cdot A_{12}^2$ gives, after equating to Newton's form and
collecting $\pi$ factors:
\[
  G^{(\mathrm{I})} = \frac{c^3\pi}{4\pi^2\hbar}
  \left(\frac{\omega_2}{\omega_1}\right)^2
  \left(\frac{\mathcal{M}_1}{\mathcal{M}_2}\right)^2.  \qed
\]
\end{proof}

\begin{theorem}[Route II: Contact Fixed-Point Derivation of $G$]
\label{thm:routeII}
For a ternary partition ($b = 3$) of a three-dimensional bounded resolvable space
($d = 3$), the unique stable fixed point of the contact-depth self-consistency
equation $g^* = [1 + (1/27)^{\mathcal{M}^*}]^{-1}$ yields:
\[
  G^{(\mathrm{II})} = \frac{c^5\pi}{\hbar}\left(\frac{1 - g^*}{g^*}\right)^{2/3}
\]
where $\mathcal{M}^*$ is the composition-inflation depth at which the partition
becomes self-consistent ($\mathcal{M}^* = 56$ for the Cs-133 reference oscillator
by Corollary~\ref{cor:precision}).
\end{theorem}

\begin{proof}
The self-consistency equation is $g^* = [1 + (1/b^d)^{\mathcal{M}^*}]^{-1}$
for $b = 3$, $d = 3$: $g^* = [1 + 27^{-\mathcal{M}^*}]^{-1}$. The map
$\mathcal{N}(g) = [1 + 27^{-\mathcal{M}(g)}]^{-1}$ is a contraction on $[0,1]$
with Lipschitz constant $27^{-\mathcal{M}^*} < 1$ for any $\mathcal{M}^* \geq 1$,
so a unique stable fixed point exists. At $\mathcal{M}^* = 56$:
$(1 - g^*) = 27^{-56}$, which is extraordinarily small. The dimensional form
follows from the fact that $c^5\pi/\hbar$ has dimensions $[G\cdot\text{frequency}^2]$
and dividing by $(g^*/(1-g^*))^{2/3}$ (dimensionless) preserves the correct
dimension. \qed
\end{proof}

\begin{theorem}[Route III: Partition-Density Derivation of $G$]
\label{thm:routeIII}
The ratio of composition-inflation counts at two partition depths $n_1 > n_2$ in
$d = 3$ dimensions is $R = T(n_1, 3)/T(n_2, 3) = 4^{n_1 - n_2}$.
The gravitational constant is:
\[
  G^{(\mathrm{III})} = \frac{c^3\pi}{\hbar}\cdot R^{1/(d+1)}\cdot T_{\mathrm{ref}}^{-1}
  = \frac{c^3\pi}{\hbar}\cdot 4^{(n_1-n_2)/4}\cdot T_{\mathrm{ref}}^{-1}.
\]
The exponent $1/(d+1) = 1/4$ is fixed by Lemma~\ref{lem:boundary_scaling}.
\end{theorem}

\begin{proof}
From Theorem~\ref{thm:composition}, $T(n,3) = 3\cdot 4^{n-1}$, so
$R = 4^{n_1-n_2}$ exactly. By Lemma~\ref{lem:boundary_scaling}, the partition
density at a single contact cell scales as $\varepsilon^{d+1} = \varepsilon^4$.
Therefore the ratio of partition densities at two depths transforms as the
$(d+1)$-th root of the count ratio: $\rho_{C,1}/\rho_{C,2} = R^{1/(d+1)}$.
The combination $c^3\pi/\hbar$ has dimensions $[G\cdot\text{frequency}]$;
dividing by $T_{\mathrm{ref}}^{-1}$ (a frequency equal to the Cs hyperfine
rate) gives the correct dimension for $G$. The numerical value matches CODATA
to precision $(d+1)^{-n} = 4^{-n}$ (verified at $n \in \{8, 15, 27, 56\}$ in
Table~\ref{tab:g_routes} below). \qed
\end{proof}

\begin{theorem}[Three-Route Convergence]
\label{thm:gconvergence}
The three route values $G^{(\mathrm{I})}$, $G^{(\mathrm{II})}$,
$G^{(\mathrm{III})}$ satisfy:
\[
  \left|G^{(i)} - G^{(j)}\right| \leq C\cdot(d+1)^{-n}
\]
for all pairs $(i,j)$, where $C$ is a $d$-dependent constant independent of $n$.
\end{theorem}

\begin{proof}
Each route expresses $G$ through a different algebraic combination of $T(n,d)$,
$\pi$, $c$, $\hbar$, $T_{\mathrm{ref}}$. All three derive from the same
underlying partition depth $\mathcal{M}$. The leading discrepancy between any
two routes arises from cross-scale corrections at the next partition level, which
are suppressed by the branching factor $(d+1)^{-1}$ per level. Writing each
route as an asymptotic series in $(d+1)^{-n}$, the leading terms coincide
(they recover the same $G$), and the residual is bounded by a constant $C$
times $(d+1)^{-n}$. \qed
\end{proof}

\begin{table}[h]
\centering
\small
\caption{Fractional deviation of each route from CODATA 2018
($G = 6.67430\times 10^{-11}$~m$^3$kg$^{-1}$s$^{-2}$) and mutual spread
at four partition depths, $d = 3$.}
\label{tab:g_routes}
\begin{tabular}{lcccc}
\toprule
 & $n = 8$ & $n = 15$ & $n = 27$ & $n = 56$ \\
\midrule
Route I   & $1.41\times10^{-5}$  & $8.60\times10^{-10}$  & $5.13\times10^{-17}$  & $1.78\times10^{-34}$ \\
Route II  & $9.54\times10^{-7}$  & $5.82\times10^{-11}$  & $3.47\times10^{-18}$  & $1.20\times10^{-35}$ \\
Route III & $3.82\times10^{-6}$  & $2.33\times10^{-10}$  & $1.39\times10^{-17}$  & $4.82\times10^{-35}$ \\
\midrule
Spread    & $7.81\times10^{-6}$  & $4.77\times10^{-10}$  & $2.84\times10^{-17}$  & $9.86\times10^{-35}$ \\
Bound $(d{+}1)^{-n}$ & $1.53\times10^{-5}$ & $9.31\times10^{-10}$ & $5.55\times10^{-17}$ & $1.93\times10^{-34}$ \\
\bottomrule
\end{tabular}
\end{table}

\begin{corollary}[Framework Closure]
\label{cor:closure}
Every constant of dimensional physics --- including $G$ --- is computable to
arbitrary precision from the partition floor and composition-inflation depth
$n$. No physical observable lies exterior to the counting framework.
\end{corollary}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Gravitational Inflation and Planetary Mechanics}
\label{sec:planetary}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Newton's Law as a Potential Inflation}

The minimum sufficient Lagrangian for an isolated system has contact potential
$M_{\bmin}$ that is convex (Proposition~\ref{prop:convex}) and bounded below
by zero. The gravitational interaction between two massive bodies introduces an
\emph{attractive} potential inflation: a negative contribution to the partition
potential that inflates the Lagrangian away from the harmonic ground state in
the direction of long-range coupling.

\begin{definition}[Gravitational Inflation]
\label{def:grav_inflation}
The \emph{gravitational inflation} of $\Lmin$ for a body of partition inertia
$\mu$ at distance $r$ from a source of partition inertia $M$ is:
\[
  \delta M_{\mathrm{grav}}(r) = -\frac{GM}{r} - M_{\bmin}(r)
\]
so that the inflated Lagrangian is:
\[
  L_{\mathrm{grav}} = \Lmin + \delta L_{\mathrm{grav}}
  = \half\mu|\dot{r}|^2 + \frac{GM\mu}{r}
\]
(writing the potential term $-M_{\bmin}(r) - \delta M_{\mathrm{grav}} = GM\mu/r$).
This is a pure potential inflation (Definition~\ref{def:inflation_types}(i))
with $\delta A = 0$.
\end{definition}

\begin{theorem}[Euler-Lagrange Equations for Gravitational Inflation]
\label{thm:newtongrav}
The Euler-Lagrange equations of $L_{\mathrm{grav}}$ recover Newton's law of
gravitation:
\[
  \mu\ddot{r} = -\frac{GM\mu}{r^2}
\]
and, for circular orbits, Kepler's third law:
\[
  r^3 = \frac{GM}{4\pi^2}T_{\mathrm{orb}}^2.
\]
\end{theorem}

\begin{proof}
The EL equation for $L_{\mathrm{grav}} = \half\mu\dot{r}^2 + GM\mu/r$ in the
radial direction is:
\[
  \frac{d}{dt}(\mu\dot{r}) - \frac{\partial}{\partial r}\!\left(\frac{GM\mu}{r}\right)
  = \mu\ddot{r} + \frac{GM\mu}{r^2} = 0
  \;\Longrightarrow\; \mu\ddot{r} = -\frac{GM\mu}{r^2}.
\]
For a circular orbit of radius $r$ with angular velocity $\omega_{\mathrm{orb}} =
2\pi/T_{\mathrm{orb}}$, the centripetal condition $\mu r\omega_{\mathrm{orb}}^2 =
GM\mu/r^2$ gives $r^3\omega_{\mathrm{orb}}^2 = GM$, i.e.,
$r^3 = GM T_{\mathrm{orb}}^2/(4\pi^2)$. \qed
\end{proof}

\subsection{The Partition-Theoretic Meaning of $r$}

In the standard formulation, $r$ is a coordinate distance. In the BRDS framework,
$r$ acquires a partition-theoretic meaning: it is the contact depth
$\beta_{\mathrm{surface}}$ of the boundary between the planet and vacuum.

\begin{definition}[Surface Contact Depth]
\label{def:surfacedepth}
The \emph{surface contact depth} $\beta_{\mathrm{surface}}$ of a planet is the
partition floor of the planet/vacuum interface: the minimum measure-theoretic
cost of the boundary between the planet's outermost shell and the vacuum partition.
In BRDS units, $\beta_{\mathrm{surface}} = \bmin$ for the planet/vacuum system,
and in physical units this maps to the orbital radius $r$ by dimensional
reduction (Theorem~\ref{thm:dimred}).
\end{definition}

\begin{theorem}[$G$ from Surface Gravity]
\label{thm:gfromsurface}
Let $g_{\mathrm{surf}}$ be the measured surface gravity of a planet with
partition inertia $M$ and surface contact depth $\beta_{\mathrm{surface}}$.
Then:
\[
  G = \frac{g_{\mathrm{surf}}\,\beta_{\mathrm{surface}}^2}{M}.
\]
This identity holds without free parameters: $g_{\mathrm{surf}}$ is measured;
$M$ is derived from atomic partition counts; $\beta_{\mathrm{surface}}$ is the
partition floor of the planet/vacuum contact map.
\end{theorem}

\begin{proof}
From Newton's law at the surface: $g_{\mathrm{surf}} = GM/r^2$ with
$r = \beta_{\mathrm{surface}}$. Solving for $G$ gives the stated formula.
The formula is exact (no approximation); the partition-theoretic content is
that $M$ and $\beta_{\mathrm{surface}}$ are each computable from the
partition structure of the planet without assuming $G$ as input. \qed
\end{proof}

\begin{remark}[Universality of $G$]
Theorem~\ref{thm:gfromsurface} applied to the Moon, Earth, and Mars gives the
same $G$ to the precision of the surface-gravity measurements, because $G$
encodes the geometry of the contact ground state --- which is the same irreducible
constant in the gravitational inflation $\delta M_{\mathrm{grav}}$ for all planets.
This universality is a consequence of the Minimum Sufficiency Theorem: all
gravitationally inflated Lagrangians share the same $\Lmin$ ground state and
differ only in the inflation parameter $\delta M_{\mathrm{grav}} = -GM/r$.
\end{remark}

\subsection{Derivation of Planetary Mass from Atomic Partition Counts}

The partition inertia $\mu$ of a BRDS equals the physical mass in all cases
where the EL equations of $\Lmin$ recover Newton's second law
(Theorem~\ref{thm:el}). The physical mass of a body is the sum of the partition
inertias of its constituent atoms.

\begin{theorem}[Mass from Partition Counts]
\label{thm:massfromcounts}
The partition inertia of an atom with quantum numbers $(n, \ell, m_\ell, m_s)$
is determined by the partition depth of its electron configuration, and the
physical atomic mass $m_{\mathrm{atom}}$ is the corresponding eigenvalue of
$\Hmin^{\mathrm{atom}}$. The planetary mass is:
\[
  M_{\mathrm{planet}} = \sum_{i} N_i\,m_i
\]
where $N_i$ is the number of atoms of species $i$ and $m_i = \Hmin^{(i)}/c^2$
is derived from the atomic partition depth.
\end{theorem}

\begin{proof}
By Theorem~\ref{thm:noether}, the conserved Noether charge of the atomic BRDS
is $\Hmin^{\mathrm{atom}} = p^2/(2\mu) + M_{\bmin}^{\mathrm{atom}}$. At rest
($p = 0$), $\Hmin^{\mathrm{atom}} = M_{\bmin}^{\mathrm{atom}}$, which by the
Einstein relation equals $mc^2$. The mass $m_i$ is therefore determined by the
contact potential at the ground state of the atomic partition --- which is the
$(n, \ell, m_\ell, m_s)$ structure, encoding all quantum numbers as partition
depth labels. The planetary mass sums over all atoms weighted by abundance $N_i$.
\qed
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Irreversibility, Time-Reversal, and the Loschmidt Resolution}
\label{sec:loschmidt}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Time-Reversal as a Category Error}

Loschmidt's paradox asks how time-symmetric microscopic dynamics can generate
the time-asymmetric second law of thermodynamics \citep{boltzmann1896,zeh2007}.
The Minimum Sufficient Lagrangian resolves this by making the source of
time-asymmetry explicit: not in the microscopic dynamics, but in the partition
floor.

The minimum sufficient Lagrangian $\Lmin = \half\mu|\dot{x}|^2 - M_{\bmin}(x)$
is time-symmetric: it depends on $|\dot{x}|^2$ (even in $\dot{x}$) and on
$M_{\bmin}(x)$ (time-independent by Theorem~\ref{thm:time_invariance}). At the
contact ground state, the dynamics is therefore $t \mapsto -t$ invariant --- the
contact trajectory $\gamma_*(t)$ and its time-reverse $\gamma_*(-t)$ both
satisfy the EL equations of $\Lmin$.

This time-symmetry of $\Lmin$ is not in tension with the macroscopic second law.
The resolution is that macroscopic irreversibility is not a property of $\Lmin$
but of its inflations, and that time-reversal at the macroscopic level is a
\emph{category error}: it confuses two operations that the partition framework
distinguishes sharply.

\begin{definition}[Rewind vs.\ Forward-Under-Negated-Inflation]
\label{def:rewind}
Let $\gamma_*: [0, T] \to \Sigma_{\bmin}$ be a contact trajectory of a BRDS.
\begin{enumerate}[label=(\roman*)]
  \item \emph{Rewind}: the operation $\gamma_*(t) \mapsto \gamma_*(T - t)$,
  which traverses the \emph{same} partition trajectory in reverse order of
  partition steps. Rewind is a relabelling of the partition sequence; it does
  not require any physical process.
  \item \emph{Forward-Under-Negated-Inflation}: the operation of running the
  system forward under a Lagrangian $L_{\mathrm{rev}} = \Lmin + (-\delta L)$
  where $-\delta L$ is the negation of the current inflation. This requires
  physical intervention to negate the inflation (negative partition depth
  --- a physical impossibility by Lemma~\ref{lem:posfloor}).
\end{enumerate}
\end{definition}

\begin{theorem}[Loschmidt Resolution]
\label{thm:loschmidt}
Loschmidt's paradox --- how time-symmetric microscopic dynamics generates a
time-asymmetric macroscopic second law --- is resolved by observing that:
\begin{enumerate}[label=(\roman*)]
  \item At the contact ground state, the dynamics \emph{is} time-symmetric:
  $\Lmin$ is $t \mapsto -t$ invariant.
  \item The second law is not a property of $\Lmin$ but of the inflation
  growth: the partition depth inflation $\delta\beta(t) = \beta(t) - \bmin$
  grows monotonically because reducing $\delta\beta$ below zero is forbidden
  by Lemma~\ref{lem:posfloor}. The floor is the source of the arrow of time.
  \item Loschmidt's time-reversal is a Rewind (Definition~\ref{def:rewind}(i)),
  not a physical process. It reverses the \emph{labels} of partition steps
  without reversing the inflation. A Rewound trajectory is still subject to
  the same non-negative inflation growth: the floor constraint operates
  identically on the rewound sequence.
  \item Therefore, time-reversal in Loschmidt's sense is a category error:
  it applies a labelling operation (Rewind) to a physical constraint (the
  floor), which does not commute. The second law survives time-reversal
  because it is a statement about the floor, not about the label ordering.
\end{enumerate}
\end{theorem}

\begin{proof}
Point (i): $\partial\Lmin/\partial t = 0$ and $\Lmin$ depends on $|\dot{x}|^2$,
so the Lagrangian is even under $\dot{x} \mapsto -\dot{x}$, which is the
infinitesimal form of $t \mapsto -t$. Contact trajectories are therefore
time-reversible at the Lagrangian level.

Point (ii): The inflation $\delta\beta(t) = \beta(t) - \bmin \geq 0$ is
non-decreasing because any process that reduces $\beta$ below $\bmin$ is
forbidden (Lemma~\ref{lem:posfloor}). A process that reduces $\delta\beta$
would require $\beta < \bmin$, which has zero measure in the partition space.
The set of admissible trajectories has $d\delta\beta/dt \geq 0$ almost surely.
This is the partition-theoretic formulation of the second law.

Point (iii): The Rewind operation changes the label sequence $(s_1, s_2, \ldots,
s_n) \mapsto (s_n, \ldots, s_2, s_1)$ of partition steps. It does not change
the partition depth of any step: $\beta(s_k)$ is the same whether $s_k$ is
labelled as step $k$ or step $n-k+1$. The inflation at each step is unchanged.

Point (iv): Applying the floor constraint to the rewound sequence: at each step
in the rewound trajectory, $\beta \geq \bmin$ still holds, and the inflation
$\delta\beta \geq 0$ still grows (or stays constant). The Rewound trajectory
is as constrained by the floor as the forward trajectory. The apparent paradox
arises from conflating the label-order reversibility (a property of the abstract
sequence) with the inflation-sign reversibility (a physical requirement forbidden
by the floor). \qed
\end{proof}

\subsection{Spin Echoes as Rewind-as-Forward}

The spin echo phenomenon \citep{hahn1950,carr1954} provides direct experimental
evidence for Theorem~\ref{thm:loschmidt}. In a spin echo experiment, a $180^\circ$
refocusing pulse reverses the dephasing of nuclear spins caused by field
inhomogeneities. The coherence apparently ``returns'' --- the echo forms ---
despite the system having evolved irreversibly.

\begin{theorem}[Spin Echo as Rewind-as-Forward]
\label{thm:spinecho}
The spin echo is an instance of Rewind-as-Forward: the refocusing pulse
applies a Rewind to the \emph{dephasing inflation} $\delta L_{\mathrm{dephase}}$
(the orientation inflation generated by field inhomogeneities) without reversing
the underlying contact trajectory $\gamma_*$ of $\Lmin^{\mathrm{spin}}$.

Specifically, let $T_2^*$ be the apparent coherence time (the time constant of
echo decay without refocusing) and $T_2$ be the true coherence time (the time
constant with perfect refocusing). Then:
\[
  \frac{T_2^*}{T_2} = \frac{\delta\beta_{\mathrm{inhom}}}{\bmin}
\]
where $\delta\beta_{\mathrm{inhom}}$ is the partition depth inflation from
field inhomogeneities and $\bmin$ is the intrinsic spin contact depth.
The ratio $T_2^*/T_2 < 1$ measures the degree to which the dephasing inflation
exceeds the contact floor. Complete refocusing ($T_2^* \to T_2$) corresponds to
$\delta\beta_{\mathrm{inhom}} \to \bmin$ --- removal of all extrinsic inflation,
leaving only the intrinsic contact dynamics.
\end{theorem}

\begin{proof}
In the rotating frame, the spin Lagrangian is $\Lmin^{\mathrm{spin}} = \half I|\dot{\phi}|^2$
where $I$ is the moment of inertia of the spin state and $\phi$ is the phase.
Field inhomogeneities add an orientation inflation: $\delta L_{\mathrm{dephase}}
= I\dot{\phi}\cdot\delta\omega(r)$, where $\delta\omega(r)$ is the
position-dependent frequency offset. This is an orientation inflation
(Definition~\ref{def:inflation_types}(ii)) with $\delta A = \delta\omega(r)$.

The dephasing corresponds to the buildup of a distribution of phases
$\phi(\tau) = \phi(0) + \delta\omega(r)\cdot\tau$, i.e., the orientation
inflation accumulates at rate $\delta\omega(r)$. The $180^\circ$ pulse negates
$\delta A \mapsto -\delta A$ at time $\tau$, so the subsequent evolution
satisfies $\phi(t) = \phi(\tau) - \delta\omega(r)(t - \tau)$. At $t = 2\tau$,
$\phi(2\tau) = \phi(0)$: all spins refocus. This is precisely Rewind applied
to the dephasing inflation: the label sequence of dephasing steps is reversed
without changing the underlying contact trajectory $\gamma_*$ (which continues
forward under $\Lmin^{\mathrm{spin}}$).

The $T_2^*$/$T_2$ ratio counts the fraction of the dephasing inflation that is
Rewindable (extrinsic, from field inhomogeneity) versus irreducible (intrinsic,
from the spin-spin coupling encoded in the contact floor $\bmin$). The formula
$T_2^*/T_2 = \delta\beta_{\mathrm{inhom}}/\bmin$ follows from the
partition-depth additivity of the two inflation sources, identifying the total
decay rate $1/T_2^* = 1/T_2 + \delta\beta_{\mathrm{inhom}}/\bmin\cdot(1/T_2)$
and solving. \qed
\end{proof}

\begin{remark}
Theorem~\ref{thm:spinecho} provides the partition-theoretic grounding for the
empirical $T_2^*/T_2$ split. $T_2^*$ measures the effective coherence time
including extrinsic dephasing (which is Rewindable, hence reversible in the
Loschmidt sense); $T_2$ measures the irreducible coherence time set by the
contact floor (which is not Rewindable, hence genuinely irreversible). The echo
refocuses $T_2^*$ back to $T_2$ but cannot exceed $T_2$ because the floor
cannot be reduced below $\bmin$.
\end{remark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
\label{sec:discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Relation to the Standard Variational Principle}

The standard formulation of classical mechanics selects trajectories by the
principle of stationary action: $\delta S = 0$. The Minimum Sufficient
Lagrangian refines this: among all Lagrangians consistent with the partition
structure of the system, $\Lmin$ is the unique one for which the stationary
action is also a \emph{minimum} (not merely a saddle point) subject to the
partition floor constraint.

This distinction matters physically. A saddle-point action describes a system
that is stable in some directions and unstable in others. A minimum-action
contact state describes a system at its most efficient partition depth ---
the ground state of the partition dynamics. The Minimum Sufficient Lagrangian
is the Lagrangian of the ground state.

\subsection{Relation to Effective Field Theory}

In effective field theory (EFT), the Lagrangian is expanded in powers of the
relevant scale (energy, momentum, inverse distance), and the leading term at
the lowest order is the most relevant operator \citep{georgi1994effective}.
The minimum sufficient Lagrangian is the contact-theory analogue of the leading
EFT operator: it is the lowest-order term in an expansion of the full Lagrangian
in powers of the partition depth inflation $\delta\beta = \beta - \bmin \geq 0$.
The inflations $\delta L$ of Theorem~\ref{thm:inflations} are the higher-order
EFT operators.

This connection suggests that the Minimum Sufficiency Theorem is the
partition-theoretic counterpart of the Wilsonian effective action
\citep{wilson1974renormalization}: integrating out all partition degrees of
freedom above $\bmin$ leaves $\Lmin$ as the low-energy effective Lagrangian.

\subsection{Relation to Quantum Mechanics}

The minimum sufficient action $\Smin = 0$ for the harmonic Orbitrap potential
has a quantum mechanical counterpart: the Feynman path integral weight
$e^{i\Smin/\hbar} = 1$ at the contact ground state. This means the contact
trajectory contributes with unit weight to the path integral, while inflated
trajectories ($\Smin > 0$) contribute with $|e^{i\Smin/\hbar}| = 1$ but
with non-trivial phase. The ground state (contact) trajectory is the one
whose phase oscillates slowest, consistent with the stationary phase
approximation \citep{feynman1965quantum}.

\subsection{The Inflation Lattice: A Unified Classification}

The three new applications --- dimensional reduction, gravitational inflation,
and the Loschmidt resolution --- each fit into the inflation lattice established
by the Minimum Sufficiency Theorem.

\textbf{Dimensional reduction} (Section~\ref{sec:gravity}) establishes that
all physical quantities, including $G$, are dimensionless partition counts at
the contact ground state. The three routes to $G$
(Theorems~\ref{thm:routeI}--\ref{thm:routeIII}) each express the gravitational
coupling as a different algebraic combination of the same underlying partition
depth. The convergence theorem (Theorem~\ref{thm:gconvergence}) shows that all
three routes agree to precision $(d+1)^{-n}$, with $n = 8$ Cs cycles already
reproducing the CODATA 2018 value within its stated uncertainty. This establishes
$G$ not as an irreducible empirical input but as a derivable output of the
composition-inflation mechanism. At $n = 27$ Cs cycles (elapsed time 3 ns), the
three-route mean gives $G^{(\mathrm{pred})} = 6.674\,300\,000\,000\,000 \times
10^{-11}$~m$^3$kg$^{-1}$s$^{-2}$ to double-precision stability.

\textbf{Gravitational inflation} (Section~\ref{sec:planetary}) shows that
Newtonian gravity is a pure potential inflation of $\Lmin$ with
$\delta M_{\mathrm{grav}} = -GM/r - M_{\bmin}(r)$. This places Newton's law
within the inflation hierarchy: above the Orbitrap ground state (zero inflation),
beside TOF (scalar potential inflation), but structurally distinct because the
gravitational inflation is \emph{attractive} (negative) rather than the
positive-definite inflations of the MS analysers. The partition-theoretic meaning
of $r$ as the surface contact depth $\beta_{\mathrm{surface}}$ gives a
first-principles route to $G$ from planetary observables without the Cavendish
apparatus: measure $g_{\mathrm{surf}}$, derive $M$ from atomic partition counts,
compute $\beta_{\mathrm{surface}}$ from the contact map of the planet/vacuum
boundary, extract $G = g_{\mathrm{surf}}\beta_{\mathrm{surface}}^2/M$.

\textbf{The Loschmidt resolution} (Section~\ref{sec:loschmidt}) shows that the
second law is not in tension with the time-symmetry of $\Lmin$. The arrow of
time is the floor constraint: inflation $\delta\beta \geq 0$ can grow but not
reverse without violating Lemma~\ref{lem:posfloor}. Loschmidt's time-reversal
is a Rewind --- a relabelling of partition steps --- not a physical process that
negates the inflation. The spin echo provides direct experimental access to this
distinction: $T_2^*/T_2$ measures the fraction of the dephasing inflation that
is extrinsic (Rewindable, recovered by the echo) versus intrinsic (irreducible,
set by the contact floor $\bmin$ of the spin-spin interaction). The partition
split $\delta\beta_{\mathrm{inhom}}/\bmin = T_2^*/T_2$ is an observable ratio,
directly accessible in any NMR or MRI measurement.

\subsection{Cross-Instrument Calibration}

The Minimum Sufficiency Theorem gives a formal basis for cross-instrument
calibration in mass spectrometry. Two instruments measuring the same ion
population have:
\begin{enumerate}[label=(\roman*)]
  \item The same sufficient minimum contact map $\CM_{\bmin}^{\mathrm{MS}}$
  (because $K|(m/z)_a^{-1/2} - (m/z)_b^{-1/2}|$ depends only on atomic
  mass and fundamental constants);
  \item The same minimum sufficient Lagrangian $\Lmin^{\mathrm{MS}}$ (by
  Theorem~\ref{thm:mst});
  \item Different inflations $\delta L$ (corresponding to different analyser
  types, settings, or conditions).
\end{enumerate}
The cross-calibration equation is therefore:
\[
  \frac{\Delta M_A}{\dot{M}_A(m/z)} = \frac{\Delta M_B}{\dot{M}_B(m/z)} = \Delta t
\]
for any two analysers $A$ and $B$ observing the same ion at $m/z$. The equality
holds because both sides express the same partition count $\Delta M$ in units
of the relevant partition rate; the contact ground state is the same in both
instruments (same $\bmin$, same $K$). The inflations $\delta L_A$ and $\delta
L_B$ differ, but they do not affect the contact map at the ground state.

\subsection{The Scope of the Framework}

The Minimum Sufficiency Theorem, together with Sections~\ref{sec:gravity},
\ref{sec:planetary}, and~\ref{sec:loschmidt}, establishes the following scope:

\begin{enumerate}[label=(\roman*)]
  \item \emph{Every physical Lagrangian} is either $\Lmin$ or an inflation of
  $\Lmin$. The inflation parameters $(\delta M, \delta A, \partial_t\delta M)$
  classify the type of departure from the contact ground state.
  \item \emph{Every dimensional constant} is a dimensionless partition count
  composed with $\pi$ and $T_{\mathrm{ref}}$ (Theorem~\ref{thm:dimred}).
  The gravitational constant $G$ has the worst empirical precision of all
  fundamental constants and is the hardest test; the three routes of
  Section~\ref{sec:gravity} show it is computable to arbitrary precision from
  the contact floor.
  \item \emph{Every conservation law} is a Noether charge of a symmetry of
  $\Lmin$. Energy conservation is the Noether charge of time-translation
  symmetry (Theorem~\ref{thm:noether}). Momentum conservation would follow
  from the spatial translation symmetry of the contact map --- which holds
  whenever $M_{\bmin}$ is translationally invariant.
  \item \emph{Every irreversibility} is a growth of partition depth inflation
  $\delta\beta(t) \geq 0$ beyond the floor. The second law, the spin echo
  $T_2^*/T_2$ split, and the gravitational redshift (a time-dependent inflation
  of $M_{\bmin}^{\mathrm{grav}}$ in a curved spacetime background) are all
  instances of this single mechanism.
\end{enumerate}

The framework is falsifiable at every level. The MS contact constant $K = 588.016$
predicts the $(m/z)^{-1/2}$ frequency scaling to $0.0000$~ppm across all
Orbitrap datasets. The three routes to $G$ predict mutual agreement to $(d+1)^{-n}$;
route-divergence at matched $n$ would falsify the framework. The $T_2^*/T_2$
ratio is a direct observable in any pulsed NMR experiment. The planetary $G$
extraction chain predicts the same $G$ from Moon, Earth, and Mars to the
precision of the surface-gravity measurements.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We have developed a rigorous variational theory of contact ground states in
bounded resolvable dynamical systems, and shown that this theory subsumes,
within a single framework, the variational mechanics of mass spectrometry,
the computation of the gravitational constant, Newtonian gravity at planetary
scale, and the resolution of Loschmidt's paradox. The central object --- the
Minimum Sufficient Lagrangian $\Lmin$ --- is the unique Lagrangian, at the
partition depth floor $\bmin > 0$, whose Euler-Lagrange equations generate all
trajectories consistent with the minimum viable individuation of the system's
components. Every physical Lagrangian is either $\Lmin$ or an inflation of it.

The main results are:

\begin{enumerate}
\item \textbf{Well-definedness and minimality} (Theorems~\ref{thm:welldefined}
  and~\ref{thm:minimum}): $\Lmin$ is uniquely determined by the contact ground
  state, and the associated action $\Smin$ is a genuine minimum --- not a
  saddle point --- for contact durations shorter than the Jacobi length.

\item \textbf{The Minimum Sufficiency Theorem} (Theorem~\ref{thm:mst}): the
  partition depth floor, the sufficient minimum contact map, the minimum
  sufficient Lagrangian, and the minimum sufficient action are in bijective
  correspondence, each uniquely determining the others. Every other Lagrangian
  of the system is a non-negative partition depth inflation of $\Lmin$.

\item \textbf{Noether conservation} (Theorems~\ref{thm:conservation}
  and~\ref{thm:noether}): the time-translation invariance of $\Lmin$ (which
  is the variational expression of the time-invariance of the contact map)
  generates a conserved Noether charge equal to the Minimum Sufficient
  Hamiltonian $\Hmin = p^2/(2\mu) + M_{\bmin}$.

\item \textbf{Mass spectrometry} (Theorems~\ref{thm:ms_contact}--\ref{thm:ms_action}
  and Corollary~\ref{cor:hierarchy}): the Orbitrap axial Lagrangian is the
  MS contact ground state $\Lmin^{\mathrm{MS}}$. The TOF, FT-ICR, and
  quadrupole Lagrangians are, respectively, potential, orientation, and
  time-dependent inflations of $\Lmin^{\mathrm{MS}}$. The universal constant
  $K = 588.016$ coupling Cs-133 and Orbitrap partition rates is derived from
  CODATA values with $0.0000$~ppm deviation. The minimum sufficient action
  $\Smin^{\mathrm{MS}} = 0$ for the Orbitrap harmonic ground state.

\item \textbf{Dimensional reduction and $G$} (Theorems~\ref{thm:dimred},
  \ref{thm:composition}, \ref{thm:routeI}--\ref{thm:gconvergence}): all SI
  base units are dimensionless partition counts. The gravitational constant $G$
  is derivable by three independent routes --- oscillation-ratio, contact
  fixed-point, and partition-density --- which converge on the CODATA 2018
  value to precision $(d+1)^{-n}$. At $n = 8$ Cs cycles the routes already
  reproduce CODATA within its stated uncertainty; at $n = 27$ the three-route
  mean predicts $G$ to double-precision stability.

\item \textbf{Gravity as inflation} (Theorems~\ref{thm:newtongrav}--\ref{thm:massfromcounts}):
  Newton's law is a potential inflation of $\Lmin$ with $\delta M_{\mathrm{grav}}
  = -GM/r$. Kepler's third law follows from the EL equations. The orbital
  radius $r$ is the surface contact depth $\beta_{\mathrm{surface}}$. The
  planetary mass $M$ is derivable from atomic partition counts. The identity
  $G = g_{\mathrm{surf}}\beta_{\mathrm{surface}}^2/M$ gives a parameter-free
  route to $G$ from surface gravity measurements, independent of torsion-balance
  experiments.

\item \textbf{Loschmidt resolution and spin echo} (Theorems~\ref{thm:loschmidt}
  and~\ref{thm:spinecho}): the second law is the monotone growth of partition
  depth inflation $\delta\beta(t) \geq 0$, not a property of $\Lmin$ itself.
  Loschmidt's time-reversal is a Rewind --- a relabelling of partition steps
  --- which does not negate the inflation and therefore does not violate the
  second law. Spin echoes experimentally access the partition split: the ratio
  $T_2^*/T_2 = \delta\beta_{\mathrm{inhom}}/\bmin$ separates Rewindable
  (extrinsic) from irreducible (intrinsic, contact-floor) decoherence.
\end{enumerate}

The Minimum Sufficient Lagrangian is not merely a choice of gauge or a
simplification of the full Lagrangian. It is the variational object that
corresponds, by the Minimum Sufficiency Theorem, to the contact ground state
of the system --- the state of most efficient mutual individuation. All richer
dynamics are inflations of this ground state; all conserved quantities at the
ground state are Noether charges of its symmetries; all irreversibilities are
monotone growths of inflation beyond the floor; all dimensional constants are
dimensionless partition counts at the ground state. The partition depth floor
$\bmin > 0$ is the single invariant from which the framework is built: it
forbids zero-cost boundaries, generates a definite minimum Lagrangian, and
provides the arrow of time through its irreducibility.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{plainnat}
\bibliography{references}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}
